---
title: '1995-10-29'
author: michael
date: '1995-10-29'
slug: '1995-10-29'
categories:
  - R
tags:
  - Lehman
  - Analyst
  - Cheung Kong
  - Hang Lung
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE}
pkgs = c("magick") 
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

# Cheung Kong meeting

1. Structure of group
1. Outlook for real estate market and comparison to press cycle
1. Presales legislation and easing impact on business
1. What is an ideal sandbank size given unattractive price outlook?
1. How to interpret falling consents to commence work ? A reflection of developer confidence in outlook for market or some other reason?
1. Net cash position ? i.e. acquisitions made over past year?

# Hang Lung meeting

1. Capital expenditure projection
1. Funding for sites recently acquired how? Fully paid?
1. Structure of group?
1. Levels of un sold inventory in market and what impact on pricing?
1. What will it take to change sentiment of consumers to consume?
1. The company mentions "trade up market" - still valid in current market conditions? 
1. Results of recent sales of a project?
1. What is status on various PRC projects? Site area, gfa, investment amount, share, cost, estimated completion date, first contribution date? 
1. Dividend policy? 
1. Major shareholder influence on business?
1. Composition of "other" revenue and operating profit? 
1. Commitments for year ahead ? Is that represented for construction spending or balance of land premium to be paid? 

# Hang Lung meeting : Amoy

1. Commentary in results suggests bearish outlook for growth ? Why?
1. One asset shows sizeable increase in rent - how much from rent rate and how much from volume ?
1. Dividends from associate? 
1. What assets acquired by company?

# Hang Lung meeting : Grand Hotels

1. Change in accounting standard for depreciation?

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "Hang Lung group schema", cache=TRUE, message=FALSE, echo=FALSE}
hl <- image_read("/Users/michaelleary/Documents/GitHub/analyst_diary/data/hanglungmindmap.jpg")
print(hl)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```

