---
title: Learning R - Rice
author: Michael
date: '2020-05-12'
slug: Rice notes
categories:
  - R
tags:
  - India
  - Agriculture
---

```{r pkgs, include=FALSE}
pkgs = c("readr", "tidyverse", "tbl2xts", "PerformanceAnalytics", "TTR", "quantmod", "tidyquant", "lubridate") 
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

## Conclusions

* Estimated farmer pretax in India of around US$84 per 20 foot container of shipped rough raw rice.

* That represents a 78% reduction in take home pretax profit since start of 2020 when estimated pretax stood at US$340/container.

* Despite some increase in price since the start of 2020 farming profitability has fallen sharply. 

### Trigger to enquiry

Article in Economic Times of India (accessed: 2020-05-12) that reported [today](https://economictimes.indiatimes.com/news/economy/agriculture/basmati-export-down-due-to-shortage-of-cargo-containers/articleshow/75689413.cms) that Basmati rice exporters are suffering from higher shipping costs as a result of lack of containers. 20 foot container rates from India to Dubai and India to Europe up around 30%. Rates to Dubai now stand at Rs 82k vs Rs60k at start of year. 

### Significance of news item 

India along with Thailand is a major exporter of rice. And rice is a significant contributor of household income in India where 1 in every 2 households are farmers. 

### Other facts

"Some 4-4.5 lakh tonne rice shipped monthly to Middle East during Ramadan. Lack of containers has halved that amount" (Economic Times article). 

1 lakh is a unit of measure of 100,000 units. So 4 lakh tonnes is 400,000 tonnes per month. 

Most shipments pass through container terminals at Mundra and Kandla [ports](https://www.google.com/maps/d/drive?state=%7B%22ids%22%3A%5B%221kiFLkKP1cLJbRZ3y_LttNOvZhnadzs6-%22%5D%2C%22action%22%3A%22open%22%2C%22userId%22%3A%22106492157068712305754%22%7D&usp=sharing) in Gujarat. 


### Container dimensions

A standard 20 foot container has volume of 33.2 m3 based on the following dimensions. 

```{r vol}
vol <- data.frame("Dimension"=c("width", "height", "length"), 
                     "external_m2" = c(2.440, 2.6, 6.058),
                     "internal_m2" = c(2.352, 2.395, 5.89))
vol
container_volume_m3 = vol[1,3]*vol[2,3]*vol[3,3]
container_volume_m3

```

### Rice volume in 20 foot container

7.2 tonnes of rice theoretically fit into a 20 foot container.

The weight of medium rough grain raw rice equates to 1 metric cup or 206g. The volume of rice in 1 cup is equivalent to 0.00095 m3. So in 1 m3 there are 216 kg of rice (taken as 206g/0.00095m3) or 1052 cups (1/0.00095) of rice.  Based on the volume of a 20 foot equivalent container of 33.2 m3 there is then potential capacity in the container of 7194 kgs or approximately 7.2 tonnes of rice. 

```{r no.tonnes}
riceweight_g = 206
cupvol_m3 = 0.00095
rice_vol_kg <- (riceweight_g/cupvol_m3)/1000
container_volume_m3
ricepercontainer_kg <- (container_volume_m3*rice_vol_kg)
no.of.tonnes<-ricepercontainer_kg/1000
no.of.tonnes

```

Based on the reported 4 lakh tonne of rice exported from India per month to the middle east (400,000 tonnes) suggests a container demand for 55,597 containers per month. 

```{r no.containers}
# calculate number of containers for export volume in tonnes
no.of.containers <- function(n) {
  n/no.of.tonnes
}
#enter volume in tonnes
no.of.containers(400000)
```

### Shipping rice cost to Dubai

Based on the reported shipping container rate of R82000/container for the India-Dubai route implies total shipping container cost monthly of Rs 4.6 billion/month or US$60 million per month. 

```{r valueDubai}
valueshipDubai <- function(p,q) {
  (((p*q)/10^6)/75)
}
#enter no of containers and rate in Rs per container
valueshipDubai(55597,82000)

```

### Rice selling price

Rice price at present is $16.8/cwt (Bloomberg).
The quantity that rice prices are quoted in is cwt that stands for 1 nr. hundred weight.

A mass measurement equal to 100 pounds of rough rice (2.22 bushels) in US definition (short hundred weight or central). In UK a 100 weight is a unit of mass equal to 112 pounds ( the long hundred weight). 

In US terms:
1 cwt = 100 lbs = 0.0453 metric tonnes = 2.2 bushels.
1 metric tonne = 2204 lbs = 22.04 cwt (hundred weight)
Therefore 1 container of 7.2 tonnes = 158.9 cwt. 

### Rice economics from Indian farmer lens

The revenue to farmer is computed as the price per hundred weight of rice by one container. At US$16.8/cwt and the container volume of 158 cwt (7.2 tonnes) and an exchange rate of 75 INR to 1 USD revenue per container in rupee is 199795.

```{r f.econ}
#revenue to farmer
no.of.tonnes
cwt_per_ton=22.04
cwt_per_container=no.of.tonnes*cwt_per_ton
rice_price_cwt = 16.8 #Rice price at present is $16.8/cwt (Bloomberg) accessed 2020-05-12
rice_rev_per_container_usd = rice_price_cwt*cwt_per_container
revenue_rp = rice_rev_per_container_usd*75
revenue_rp
```

Based on 2014/15 production [cost](http://agritech.tnau.ac.in/agriculture/agri_costofcultivation_rice.html) per quintal of rice of Rs 1549 total production costs total Rp 111443. A quintal is taken as 100 kg. So one container of 7.2 tonnes is around 72 quintals. To these costs are added shipping cost of Rs 82000 per container. 

```{r f.costs}
#costs borne by farmer
cost_per_quintal_rp=1549
quintal_in_kg = 100
no_of_quintal_per_container=ricepercontainer_kg/quintal_in_kg
rice_cost_rp = no_of_quintal_per_container*cost_per_quintal_rp
shipping_cost=82000
localcosts = rice_cost_rp+shipping_cost
localcosts
```

The result of subtracting revenue per container from its associated costs of production and shipping costs is US$84. That is based on current shipping rate and the current price of rice. 

```{r f.pretax}
#farming rice pretax profit
pretax_rp = revenue_rp-localcosts
pretax_rp
```

```{r pretaxusd}
#US 
pretax_usd = pretax_rp/75
pretax_usd
```

### Historic rice prices 

Further examination is made on the trend in rice price over time to better appreciate farmer profitability at the top line. 


Rice price data is quoted in two different ways. Thailand rice prices which are quoted in US$/metric tonne ie. by the St Louis Federal Reserve of [FRED](https://fred.stlouisfed.org/series/PRICENPQUSDM) that shows a March 2020 Thai rice price at US$470/tonne. And prices quoted by Bloomberg that are in US$/cwt. 

Given the ratio of 1 tonne to 22 cwt the Thai price is equivalent to US$21/cwt (470/22) that compares to the Bloomberg rice price of US$16.80/cwt. 



```{r rp, include=FALSE}
### read in data
ricep <- read_csv("~/Downloads/roughrice.csv")

### view data as tibble df

ricep

###  confirm time variabe is present in tibble (TRUE confirms) 


ricep[,"date"] %>% .[[1]] %>% timeBased()

###  set variable to represent data set

tbldata <- ricep

###  convert dataset from tibble to xts

datats <- xtsdata <- ricep %>% 
  tbl_xts(., cols_to_xts = "value")

###  review structure of data set

str(datats)
class(datats)

###  to transform back to tibble object

# datats %>% xts_tbl()
#str(datats)

###  To generate metrics

tbldata %>% 
  tbl_xts(., cols_to_xts = "value") %>% 
  lapply(.,Return.calculate, "discrete") %>% Reduce(merge,.) %>% table.DownsideRisk(.)

###  Yearly returns

yearrtn <- ricep %>% tbl_xts(.) %>%  
  PerformanceAnalytics::Return.calculate(.) %>% 
  xts::apply.yearly(., FUN = PerformanceAnalytics::StdDev.annualized) %>% 
  xts_tbl

tail(yearrtn,14)


###  Chart of CVaR

ricep %>% tbl_xts(.) %>%  
  PerformanceAnalytics::Return.calculate(.) %>% 
  xts::apply.yearly(., FUN = PerformanceAnalytics::CVaR) %>% 
  xts_tbl %>% tidyr::gather(value, CVaR, -date) %>% 
  ggplot() + geom_line(aes(date, CVaR)) + theme_bw() + facet_wrap(~value)


###  Time plot hisory using XTS

# glimpse(datats)
#head(datats)

# plot the Close
#plot(datats[,"value"])

# plot a subset of the data
#plot(datats[,"value"], subset="2020-01-01/2020-05-01")

#simple.ret <- function(x, col.name){
#  x[,col.name] / lag(x[,col.name]) - 1
#}

# pass in the simple.ret function and plot the output
#plot(datats, FUN=simple.ret, col.name="value")

# plot the close and add a panel with the simple returns
#plot(datats[,"value"])
#R <- simple.ret(datats, "value")
#lines(R, type="h")

# add the 200 period simple moving average to panel 1 of the plot
#plot(datats[,"value"], main="CBOT rough rice US$/cwt")
#lines(SMA(datats[,"value"], n = 200), on=1, col="blue")


# add year end points to the chart
#points(datats[endpoints(datats[,"value"], on = "years"), "value"], 
#       col="red", pch=17, on=1)

# add legend to panel 1
#addLegend("topright", on=1, 
#          legend.names = c("value", "SMA(200)"), 
#          lty=c(1, 1), lwd=c(2, 1),
#          col=c("black", "blue", "red"))


# Time plot 2020

#glimpse(datats)
#head(datats)
#x <- datats['2019/2020'] # All of 2019 and 2020

# plot the Close
#plot(x[,"value"])

# plot a subset of the data
#x <- plot(x[,"value"], subset="2019-01-01/2020-05-01")

#simple.ret <- function(x, col.name){
#  x[,col.name] / lag(x[,col.name]) - 1
#}

# pass in the simple.ret function and plot the output
#plot(x, FUN=simple.ret, col.name="value")

# plot the close and add a panel with the simple returns
#plot(x[,"value"])
#R <- simple.ret(x, "value")
#lines(R, type="h")

# add the 50 period simple moving average to panel 1 of the plot
#plot(x[,"value"], 
#     main="CBOT rough rice US$/cwt")
#    lines(SMA(x[,"value"], n = 50), on=1, col="blue")

# add month end points to the chart
#points(x[endpoints(x[,"value"], on = "months"), "value"], 
#       col="red", pch=17, on=1)

# add legend to panel 1
#addLegend("topleft", on=1, 
#          legend.names = c("value", "SMA(50)"), 
#          lty=c(1, 1), lwd=c(2, 1),
#          col=c("black", "blue", "red"))

```

###  Plotting with tidyquant

###  add the 200 period simple moving average to panel 1 of the plot

```{r rp20}
ricep %>% 
  ggplot(aes(x=date, y=value))+
  geom_line() +
  labs(title="CBOT Rough Rice price", y="US$.cwt", x="") +
  theme_tq()
```

###  focus on 2019-2020 data

```{r ricepr, include=FALSE}
ricep %>% 
  ggplot(aes(x=date, y=value))+
  geom_line() +
  labs(title="CBOT Rough Rice price", 
       subtitle="2019-2020 trend",
       y="US$.cwt", x="") +
  coord_x_date(xlim=c("2019-01-01", "2020--05-10"),
                ylim=c(5,20)) +
  theme_tq()
```

###  add 50 day simple moving average

```{r 50daysma}
#mytheme <- theme_grey()+theme(legend.position="right")
#theme_set(mytheme)

ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line() +
    geom_ma(ma_fun = SMA, n = 50, color = 'blue', linetype = 3, size = 1) +
    geom_ma(ma_fun = SMA, n = 200, color = "red", linetype= 5, size = 1) +
  scale_color_manual(name="", values=c("Important line"='grey', "50ma"='red', "200ms"='green'))+
  labs(title="CBOT Rough Rice price", 
       subtitle="50 and 200 day SMA",
       y="US$.cwt", x="") +
  coord_x_date(xlim=c("2019-01-01", "2020--05-10"),
                ylim=c(5,20))+
  theme_tq()

```

```{r legendricepr,  echo=FALSE, Include=FALSE}

# test plot with legend
#https://ggplot2-book.org/scales.html
df <- data.frame(x = 1:3, y = 1:3, z = c("a", "b", "c"))
base <- ggplot(df, aes(x, y)) + 
  geom_point(aes(colour = z), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)
```

```{r echo=FALSE, Include=FALSE}
#base + theme(legend.position = "right") # the default 
#base + theme(legend.position = "bottom")
#base + theme(legend.position = "none")
```

###  add exponential moving average

```{r expmovgavg}
ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line() +
    geom_ma(ma_fun = EMA, n = 50, linetype = 5, size = 1.25) +
    geom_ma(ma_fun = EMA, n = 200, color = "red", size = 1.25) + 
    labs(title="CBOT Rough Rice price", 
       subtitle="50 and 200 day EMA",
       y="US$.cwt", x="") +
  coord_x_date(xlim=c("2019-01-01", "2020--05-10"),
                ylim=c(5,20)) +
  theme_tq()
```

###  continuous scale

```{r continu}
ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line(color = palette()[[1]]) +
  scale_y_continuous() +
  labs(title="CBOT Rough Rice price", 
       subtitle="Continuous scale",
       y="US$.cwt", x="") +
  theme_tq()

```

###  Log scale

```{r logscalericep}
ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line(color = palette()[[1]]) +
  scale_y_log10() +
  labs(title="CBOT Rough Rice price", 
       subtitle="Log scale",
       y="US$.cwt", x="") +
  theme_tq()

```

###  Regression trend lines

```{r trendlineprice}
ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line(color = palette()[[1]]) +
  scale_y_log10() +
  geom_smooth(method="lm")+
  labs(title="CBOT Rough Rice price", 
       subtitle="Log scale, Applying linear trend line",
       y="US$.cwt", x="") +
  theme_tq()
```

```{r echo=FALSE, include=FALSE}
###  Themes
n_200 <- 200 # Number of periods (days) for moving average
n_50 <- 50 # Number of periods (days) for moving average
ricep %>%
    ggplot(aes(x = date, y = value)) +
    geom_line(size = 1) +
  geom_ma(n=15, color="darkblue", size=1)+
  geom_ma(n=n_50, color="green", size=0.5)+
  geom_ma(n=n_200, color="red", size=1)+
    labs(title = "CBOT Rough Rice price", 
       y = "US$/cwt")+
  coord_x_date(xlim = c("2007-01-01", "2020-05-10"))+
    theme_tq_dark() +
    scale_color_tq(theme = "dark") +
    scale_y_continuous(labels = scales::dollar)
```





