---
title: Highways in India
author: michael
date: '2020-07-06'
slug: highways-in-india
categories:
  - R
tags:
  - Highways
  - Transport
  - Communication
  - India
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE}
pkgs = c("dplyr","ggplot2", "kableExtra", "plyr", "cowplot") 
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

```{r, include=FALSE}
#x <- readRDS("/Users/michaelleary/testR/x.rds") #in rds form
roads <- read.csv(file = "~/Documents/GitHub/IndiaRoads/indhighway.csv", stringsAsFactors = TRUE) #in csv form

```

```{r, include=FALSE}
str(roads)
```

```{r, include=FALSE}
# transform factor to numeric values 
roads$highwaycode <- as.factor(roads$highwaycode)
```

```{r, include=FALSE}
str(roads)
```

```{r, include=FALSE}
ds <- roads
```

```{r, include=FALSE}
dsmat <- as.matrix(ds$kms_in_state)
totallength <- colSums(dsmat)

```

```{r, include=FALSE}
summary(ds)
```

```{r, include=FALSE}
attributes(ds)
```

```{r, include=FALSE}
ds["state_passing"]
```

```{r, include=FALSE}
ds[ds$state_passing=="Assam",]
```

```{r, include=FALSE}
ds %>%
  filter((state_passing == "Maharashtra") ) %>%
  select(state_passing, new_nh_no, kms_in_state, highwaycode)
  
```

```{r, include=FALSE}
ds %>%
  group_by(state_passing, highwaycode) %>%
  summarise_at("kms_in_state", sum)
```

```{r, include=FALSE}
dso <- ds[order(ds$state_passing),]

plot(ds$state_passing, ds$kms_in_state)
```

```{r, include=FALSE}
ggplot(data=dso, aes(y=ds$state_passing)) + geom_bar(stat="count")
```

```{r, include=FALSE}
plot(ds)
plot(as.numeric(ds$state_passing), ds$kms_in_state)
qplot(y=ds$state_passing, ds$kms_in_state)
```

```{r, include=FALSE}
qplot(highwaycode, kms_in_state, data = ds, geom="boxplot", fill=state_passing)
```

```{r, include=FALSE}
mu <- ddply(ds, "highwaycode", summarise, grp.mean=mean(kms_in_state))
mu
```

```{r, include=FALSE}
skimr::skim(ds)
```

```{r, include=FALSE}
ds %>%
  group_by(state_passing, highwaycode) %>%
  dplyr::summarise(count=n())
```

```{r, include=FALSE}
ds %>%
  group_by(state_passing, highwaycode) %>%
  dplyr::summarise(count=n(),
                   sum = sum(kms_in_state))
```

## Task

This task focused on India's national highway system. Two tables are generated that show the number of states each highway passes through and the length of each highway. 

Data was sourced from [Ministry of Road Transport and Highways's](https://morth.nic.in/sites/default/files/Details-of-National-Highways-as-on-31.03_1.pdf) report dated 2019-03-31 (accessed on 2020-07-06).  

The data set contained `r nrow(ds)` observations of India highways for `r totallength` kilometers. Each row gives the kilometers of highway in each state it traverses. A factor was added to the data set to distinguish between national highways that have a numeric label and highways with a non numeric label. 

#### Other links

1. 2010 List of highways - [Wikipedia](https://en.wikipedia.org/wiki/List_of_National_Highways_in_India_by_highway_number)

1. Highways map - [Wikipedia](https://en.wikipedia.org/wiki/List_of_National_Highways_in_India_by_highway_number#/media/File:Renumbered_National_Highways_map_of_India_(Schematic).jpg)




```{r, include=FALSE}
statehighwaysummary <- ds %>%
  group_by(state_passing) %>%
  dplyr::summarise(count=n(),
                   sum_kms = sum(kms_in_state))
```

```{r, include=FALSE}
statehighwaysummary
```

\newpage

Count and length of national highways by State

```{r, echo=FALSE}
statehighwaysummary %>%
  kable(caption="National highways by state by kms") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r, include=FALSE}
statehighwaylength <- ds %>%
  group_by(new_nh_no) %>%
  dplyr::summarise(count=n(),
                   sum_kms = sum(kms_in_state)) %>%
  arrange(desc(sum_kms))
```

```{r kablefunction, include=FALSE}
my_kable = function(x, max.rows=20, ...) {
  kable(x[1:max.rows, ], ...)
}
```

\newpage

Count of states traversed and length of national highways 


```{r, echo=FALSE}
statehighwaylength %>%
  my_kable( caption="20 longest national highways in India") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

### Highway links

* [nh-27](https://en.wikipedia.org/wiki/National_Highway_27_(India))
* [nh-44](https://en.wikipedia.org/wiki/National_Highway_44_(India))
* [nh-48](https://en.wikipedia.org/wiki/National_Highway_48_(India))
* [nh-52](https://en.wikipedia.org/wiki/National_Highway_52_(India))
* [nh-30](https://en.wikipedia.org/wiki/National_Highway_30_(India))










