---
title: 'Company results template: share price'
author: "Michael"
date: '2020-05-27'
slug: Company results template
categories: R
tags:
- Finance
- Investment research
- Tools
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

```{r packages, include=FALSE}
# Packages used
pkgs = c("tidyverse", 
         "tidyquant", 
         "magrittr", 
         "data.table", 
         "fBasics", 
         "gdata",
         "dplyr",
         "ggplot2",
         "spelling",
         "tufte") 
#package names: 
#tidyverse for operations using dplyr's piping feature; 
#magrittr %<>% function pipes left hand data to right side and returns result to left hand overwriting original contents on left hand
#data.table for fast csv save
#gdata has keep function for permitting convenient trim of env objects with sure=TRUE

install.packages(pkgs, repos="https://cran.r-project.org")

inst = lapply(pkgs, library, character.only = TRUE) #load packages

```

```{r, include = FALSE}
symbol <- c('7269.T', '^N225', '^DJI')
company <- c("Suzuki", "N225 Index","Dow Jones Index")


#Map a tibble with selected variables companies and index items
stock_list <- tibble(symbol = symbol,
                     company = company)
# companies excluding indexes
stock.list.co <- filter(stock_list, 
                    c(!(symbol=="^DJI" | symbol=="^N225" )))

```

```{r, echo = FALSE}

#extract prices
p <- stock_list %>% tq_get(get = "stock.prices", from = "2020-01-01", to = Sys.Date())

# In tibble data frame price data for company is stacked on each other.  
```

```{r organizedata, include = FALSE}

motor.sp <- p

#list column names
colnames <- colnames(motor.sp)
names(motor.sp)[3] <- "t"

# prepare to convert character variables symbol, company to factors

motor.spf <- motor.sp

# save character variables
motor.spf %>%
sapply(is.character) %>%
which() %>%
print() ->
chari

# alternate save of character variable names
motor.spf %>%
names() %>%
'['(chari) %>%
print() ->
charc

# convert character variables to factors
motor.spf[charc] %<>% map(factor)  #uncheck
str(motor.spf)
glimpse(motor.spf)

```

# Purpose

Thought process and actions steps following on earnings announcement. The example am using is the fourth quarter results posted by Suzuki Motors (7269.T, JPY3700) for period ended 2020-03-31.

## Recent price performance

Recent price analysis versus benchmarks show the shares of Suzuki remain 26% below the peak in the year and below that of the Nikkei and Dow Jones.   

```{r summarytable, echo=FALSE}
sum.motor <- motor.spf %>%
  group_by(symbol, company) %>%
  summarize(curr_price = last(adjusted), 
            min_price = round(min(adjusted),2), 
            max_price = round(max(adjusted),2),
            pctchange_from_max = round(100 * (curr_price-max_price)/max_price)) %>%
  arrange(desc(pctchange_from_max))

sum.motor

```

```{r shareplot, echo=FALSE}

# https://ggplot2-book.org/facet.html
adj.plot <- ggplot(data=motor.spf, 
      aes(t, adjusted))
adj.plot + geom_line() + facet_wrap(~symbol, scales = "free")

```
```{r 5dayperf, echo=FALSE, include=FALSE}

head(motor.spf)
str(motor.spf)

motor.spf %>%
  dplyr::filter(t == c("2020-05-20")) %>%
  select(symbol, t, adjusted)

todate <- Sys.Date()-5
t5day <- seq(Sys.Date(), todate, by = -1)

motor.spf %>%
  select(symbol, t, adjusted) %>%
  dplyr::filter(symbol == "7269.T",
                t >= "2020-05-20")

# date vector of every date betwee these two date and includes days of no shares
datevector <- seq(Sys.Date(), Sys.Date()-6, by=-1)


sh5day <- motor.spf %>%
  select(symbol, t, volume, adjusted) %>%
  dplyr::filter(symbol == "7269.T",
                t >= Sys.Date()-6)

```

Ahead of the results the shares have picked up modestly. 

```{r dh5day, echo=FALSE}
sh5day
```

That price and volume performance compares with average, median, sd, min and max since start of year. 

```{r sumdji, include=FALSE}

# summary values for grouped symbols

mean_a <- tapply(motor.spf$adjusted, motor.spf$symbol, mean)
median_b <- tapply(motor.spf$adjusted, motor.spf$symbol, median)
sd_c <- tapply(motor.spf$adjusted, motor.spf$symbol, sd)
min_d <- tapply(motor.spf$adjusted, motor.spf$symbol, min)
max_e <- tapply(motor.spf$adjusted, motor.spf$symbol, max)

f <- rbind(mean_a,median_b,sd_c,min_d,max_e)

```

```{r ytdperf, echo=FALSE}
f
```

```{r relperffullperiod, echo=FALSE, include=FALSE}

ds1 <- motor.spf
stk <- ds1 %>% dplyr::filter((symbol=="7269.T") )
stkn225 <- ds1 %>% dplyr::filter((symbol=="^N225") )
stkdji <- ds1 %>% dplyr::filter((symbol=="^DJI") )

head(stk)
teststrip2 <- stk[,-c(1:2, 4:8)]
stka <- rename(teststrip2, sa=adjusted)

head(stkn225)
teststrip2 <- stkn225[,-c(1:2, 4:8)]
stkn <- rename(teststrip2, n225a=adjusted)

head(stkdji)
teststrip2 <- stkdji[,-c(1:2, 4:8)]
stkd <- rename(teststrip2, djia=adjusted)

#https://jozef.io/r006-merge/
# merge tibbles on date column

dst <- dplyr::full_join(stka, stkn, by="t")
head(dst)
dstall <- dplyr::full_join(dst, stkd, by="t")
head(dstall)
d <- dstall
head(d)

dst <- transform(d, sa.idx=100*d$sa/d$sa[1], n.idx=100*d$n225a/d$n225a[1], d.idx=100*d$djia/d$djia[1])
head(dst)

```

```{r relperforchart, echo=FALSE, include=FALSE}
basketplot <- ggplot(dst, aes(t)) +
  geom_line(aes(y=sa.idx, color="Suzuki")) + 
  geom_line(aes(y=n.idx, color="N225")) +
  geom_line(aes(y=d.idx, color="DJIA")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Suzuki, N225 and DJIA index", x="Months", y="Index", color="legend") +
  theme_minimal()
  
```

With the price of the shares and the Dow Jones and Nikkei indexed to 100 at the start of the year the shares have lagged.  

```{r basketplot, echo=FALSE, warning=FALSE}
plot(basketplot)
```