---
title: Column add and reorder 
author: michael
date: '2020-06-01'
slug: working-with-dataset
categories:
  - R
tags:
  - Tools
---

## Task:
+ Extract table data from [Wikipedia](https://en.wikipedia.org/wiki/List_of_skeletal_muscles_of_the_human_body);
+ Combine table rows into one data frame;
+ Add a column to combined data frame that provides name of original dataframe - so as to facilitate identifying source of data
+ Reorder the columns in expanded data frame to put orgin data frame column first. 

There are 48 tables of muscle groups in the human body listed on the wikipedia page. Each table was extracted and stored in a separate data frame. A comma separated list of individual data frames was put into a vector. And a function then appended a column with the name of individual source data frame to each row of data. After that the individual data frames were bounded into one object. After the appended column and merged data frame operations was complete the variable columns were reordered. 

### Future
add a fresh column with recategorised groups

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
pkgs = c("rvest", "tidyverse", "janitor", "lubridate", "dplyr", "kableExtra", "data.table", "forcats") 
# lubridate for %m+% function
# rvest for web scraping
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

```{r datevars, include=FALSE}
# save current system date to a variable
today <- Sys.Date()
# subtract 3 months from the current date
date <- today %m+% months(-3)
print(date)
```



```{r rawdatacollect, include=FALSE}
# get the URL for the wikipedia page with all states
url <- "https://en.wikipedia.org/wiki/List_of_skeletal_muscles_of_the_human_body"

# use that URL to scrape the states table using rvest
forehead.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[1]') %>% 
  html_table(fill=TRUE)
#create a vector of states
forehead <- forehead.m[[1]]

extraocular.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[2]') %>% 
  html_table(fill=TRUE)
#create a vector of states
extraocular <- extraocular.m[[1]]

ear.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[3]') %>% 
  html_table(fill=TRUE)
#create a vector of states
ear <- ear.m[[1]]

nose.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[4]') %>% 
  html_table(fill=TRUE)
#create a vector of states
nose <- nose.m[[1]]

mouth.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[5]') %>% 
  html_table(fill=TRUE)
#create a vector of states
mouth <- mouth.m[[1]]

mastication.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[6]') %>% 
  html_table(fill=TRUE)
#create a vector of states
mastication <- mastication.m[[1]]

tongueext.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[7]') %>% 
  html_table(fill=TRUE)
#create a vector of states
tongueextrinsic <- tongueext.m[[1]]

tongueint.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[8]') %>% 
  html_table(fill=TRUE)
#create a vector of states
tongueintrinsic <- tongueint.m[[1]]

softpalate.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[9]') %>% 
  html_table(fill=TRUE)
#create a vector of states
softpalate <- softpalate.m[[1]]

pharynx.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[10]') %>% 
  html_table(fill=TRUE)
#create a vector of states
pharynx <- pharynx.m[[1]]

larynx.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[11]') %>% 
  html_table(fill=TRUE)
#create a vector of states
larynx <- larynx.m[[1]]

neck.clavicular.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[12]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.clavicular <- neck.clavicular.m[[1]]

neck.suprahyoid.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[13]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.suprahyoid <- neck.suprahyoid.m[[1]]

neck.infrahyoid.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[14]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.infrahyoid <- neck.infrahyoid.m[[1]]

neck.anterior.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[15]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.anterior <- neck.anterior.m[[1]]

neck.lateral.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[16]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.lateral <- neck.lateral.m[[1]]

neck.posterior.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[17]') %>% 
  html_table(fill=TRUE)
#create a vector of states
neck.posterior <- neck.posterior.m[[1]]

torso.back.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[18]') %>% 
  html_table(fill=TRUE)
#create a vector of states
torso.back<- torso.back.m[[1]]

chest.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[19]') %>% 
  html_table(fill=TRUE)
#create a vector of states
chest<- chest.m[[1]]

abdomen.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[20]') %>% 
  html_table(fill=TRUE)
#create a vector of states
abdomen<- abdomen.m[[1]]

pelvis.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[21]') %>% 
  html_table(fill=TRUE)
#create a vector of states
pelvis<- pelvis.m[[1]]

perineum.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[22]') %>% 
  html_table(fill=TRUE)
#create a vector of states
perineum<- perineum.m[[1]]

upperlimb.vertebralcol.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[23]') %>% 
  html_table(fill=TRUE)
#create a vector of states
upperlimb.vertebralcol<- upperlimb.vertebralcol.m[[1]]

upperlimb.thoracicwall.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[24]') %>% 
  html_table(fill=TRUE)
#create a vector of states
upperlimb.thoracicwall<- upperlimb.thoracicwall.m[[1]]

upperlimb.shoulder.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[25]') %>% 
  html_table(fill=TRUE)
#create a vector of states
upperlimb.shoulder<- upperlimb.shoulder.m[[1]]

upperlimb.armanterior.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[26]') %>% 
  html_table(fill=TRUE)
#create a vector of states
upperlimbarmanterior.arm<- upperlimb.armanterior.m[[1]]

upperlimb.armposterior.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[27]') %>% 
  html_table(fill=TRUE)
#create a vector of states
upperlimbarmposterior.arm<- upperlimb.armposterior.m[[1]]

forearm.anteriorcompartment.superficial.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[28]') %>% 
  html_table(fill=TRUE)
#create a vector of states
forearmanteriorcompartment.superficial <- forearm.anteriorcompartment.superficial.m[[1]]

forearm.anteriorcompartment.deep.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[29]') %>% 
  html_table(fill=TRUE)
#create a vector of states
forearmanteriorcompartment.deep <- forearm.anteriorcompartment.deep.m[[1]]

forearm.posteriorcompartment.superficial.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[30]') %>% 
  html_table(fill=TRUE)
#create a vector of states
forearmposteriorcompartment.superficial <- forearm.posteriorcompartment.superficial.m[[1]]

forearm.posteriorcompartment.deep.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[31]') %>% 
  html_table(fill=TRUE)
#create a vector of states
forearmposteriorcompartment.deep <- forearm.posteriorcompartment.deep.m[[1]]

handlateralcolarthenar.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[32]') %>% 
  html_table(fill=TRUE)
#create a vector of states
handlateralcolarthenar <- handlateralcolarthenar.m[[1]]

handmedialvolar.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[33]') %>% 
  html_table(fill=TRUE)
#create a vector of states
handmedialvolar <- handmedialvolar.m[[1]]

handintermediate.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[34]') %>% 
  html_table(fill=TRUE)
#create a vector of states
handintermediate <- handintermediate.m[[1]]

loweriliac.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[35]') %>% 
  html_table(fill=TRUE)
#create a vector of states
loweriliac <- loweriliac.m[[1]]

lowergluteal.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[36]') %>% 
  html_table(fill=TRUE)
#create a vector of states
lowergluteal <- lowergluteal.m[[1]]

thighanteriorcompartment.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[37]') %>% 
  html_table(fill=TRUE)
#create a vector of states
thighanteriorcompartment <- thighanteriorcompartment.m[[1]]

thighposteriorcompartment.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[38]') %>% 
  html_table(fill=TRUE)
#create a vector of states
thighposteriorcompartment <-thighposteriorcompartment.m[[1]]

thighmedialcompartment.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[39]') %>% 
  html_table(fill=TRUE)
#create a vector of states
thighmedialcompartment <-thighmedialcompartment.m[[1]]

leganteriorcompartment.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[40]') %>% 
  html_table(fill=TRUE)
#create a vector of states
leganteriorcompartment <- leganteriorcompartment.m[[1]]

legposteriorcompartmentsuperficial.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[41]') %>% 
  html_table(fill=TRUE)
#create a vector of states
legposteriorcompartmentsuperficial <- legposteriorcompartmentsuperficial.m[[1]]

legposteriorcompartmentdeep.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[42]') %>% 
  html_table(fill=TRUE)
#create a vector of states
legposteriorcompartmentdeep <- legposteriorcompartmentdeep.m[[1]]

leglateralcompartmentdeep.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[43]') %>% 
  html_table(fill=TRUE)
#create a vector of states
leglateralcompartmentdeep <- leglateralcompartmentdeep.m[[1]]

footdorsal.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[44]') %>% 
  html_table(fill=TRUE)
#create a vector of states
footdorsal <- footdorsal.m[[1]]

footplantarfirstlayer.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[45]') %>% 
  html_table(fill=TRUE)
#create a vector of states
footplantarfirstlayer <- footplantarfirstlayer.m[[1]]

footplantarsecondlayer.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[46]') %>% 
  html_table(fill=TRUE)
#create a vector of states
footplantarsecondlayer <- footplantarsecondlayer.m[[1]]

footplantarthirdlayer.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[47]') %>% 
  html_table(fill=TRUE)
#create a vector of states
footplantarthirdlayer <- footplantarthirdlayer.m[[1]]

footplantarfourthlayer.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table[48]') %>% 
  html_table(fill=TRUE)
#create a vector of states
footplantarfourthlayer <- footplantarfourthlayer.m[[1]]


```



```{r, include=FALSE}
# add column for variable name in each df

ID <- c("forehead",
"extraocular",
"ear",
"nose",
"mouth",
"mastication",
"tongueextrinsic",
"tongueintrinsic",
"softpalate",
"pharynx",
"larynx",
"neck.clavicular",
"neck.suprahyoid",
"neck.infrahyoid",
"neck.anterior",
"neck.lateral",
"neck.posterior",
"torso.back",
"chest",
"abdomen",
"pelvis",
"perineum",
"upperlimb.vertebralcol",
"upperlimb.thoracicwall",
"upperlimb.shoulder",
"upperlimbarmanterior.arm",
"upperlimbarmposterior.arm",
"forearmanteriorcompartment.superficial",
"forearmanteriorcompartment.deep",
"forearmposteriorcompartment.superficial",
"forearmposteriorcompartment.deep",
"handlateralcolarthenar",
"handmedialvolar",
"handintermediate",
"loweriliac",
"lowergluteal",
"thighanteriorcompartment",
"thighposteriorcompartment",
"thighmedialcompartment",
"leganteriorcompartment",
"legposteriorcompartmentsuperficial",
"legposteriorcompartmentdeep",
"leglateralcompartmentdeep",
"footdorsal",
"footplantarfirstlayer",
"footplantarsecondlayer",
"footplantarthirdlayer",
"footplantarfourthlayer")

```

```{r appendfunction, include=FALSE}

# Function to combine (rbind) data frames and create column with name of original data frames

# Source: https://stackoverflow.com/questions/15162197/combine-rbind-data-frames-and-create-column-with-name-of-original-data-frames

AppendMe <- function(dfNames) {
  do.call(rbind, lapply(dfNames, function(x) {
    cbind(get(x), source = x)
  }))
}
```

```{r, include=FALSE }
# append column to dataset with name of source variable in rbind operations
bodywhole2 <- AppendMe(c(ID))
```


```{r, include=FALSE}
# get column names
# source: http://www.sthda.com/english/wiki/reordering-data-frame-columns-in-r
colnames(bodywhole2)

```

```{r, include=FALSE}
bodywhole3 <- bodywhole2[, c(8, 1, 2, 3,4,5,6,7)]
tail(bodywhole3)
```

\newpage

```{r arrangedtable, include=FALSE}
# source: https://chartio.com/resources/tutorials/how-to-sort-a-data-frame-by-multiple-columns-in-r/

# first sort by source and then by muscle

tab <- bodywhole3 [
  order( bodywhole3[,1], bodywhole3[,2]),
]

```

```{r tabtable, echo=FALSE, warnings=FALSE}
knitr::kable(tab[,c(1,2)], "html",
             caption = "Body muscles",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

