---
title: Asset appraisal
subtitle: "investment committee memo"
author: michael
date: '2020-06-07'
slug: asset-appraisal
categories:
  - R
tags:
  - Tools
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Consider

What we want to know in evaluation of investments. 

1. What has happened over the past two years?
1. What has been purchased?  
1. Why were the assets purchased, purchased?
1. How have they been doing?
1. Why were they purchased that way?
1. How do you make the purchases happen?
1. What are the plans for the future?
1. Why have these plans?
1. What do you see for the future?
1. How do we know for sure?
1. What is the risk?
1. How do we plan to make money?
1. How do we protect our investment? Who/why
1. What do you want? What is memo asking for? 
1. How do you justify the request? 


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```