---
title: Bihar background
author: michael
date: '2020-06-30'
slug: bihar-background
categories:
  - R
tags:
  - Governance
  - Politics
  - India
  - Bihar
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r pkgsused, include=FALSE}
pkgs = c("magick", "rsvg", "rvest", "kableExtra", "dplyr", "janitor", "readr") 
# manage images
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages

```

## Task

Aggregate data on Bihar

## Data

Sourced from Wikipedia and Wikimedia commons 

## Electoral calendar

```{r data, include=FALSE}
cal <- read_csv("/Users/michaelleary/Documents/GitHub/generalitems/biharcal.csv", col_names = TRUE,)
```

```{r biharcal, include=FALSE}
cal_df <- as.data.frame(cal)
cal_df
```

```{r biharcalendar, echo=FALSE}
knitr::kable(cal_df[,c(1:4)], "html",
             caption = "Election calendar with focus on Bihar",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

## Other information 

R packages used include magick^[Jeroen Ooms (2020). magick: Advanced Graphics and Image-Processing in R. R package version 2.4.0. https://CRAN.R-project.org/package=magick], rsvg^[Jeroen Ooms (2020). rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays. R package version 2.1. https://CRAN.R-project.org/package=rsvg], rvest^[Hadley Wickham (2019). rvest: Easily Harvest (Scrape) Web Pages. R package version 0.3.5. https://CRAN.R-project.org/package=rvest], and, kableExtra^[Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R
  package version 1.1.0. https://CRAN.R-project.org/package=kableExtra ]. 

\newpage

## 1. State seal 

* Original file size: 458x548 pixels and 18KB. Image source: [Fred the Oyster](https://commons.wikimedia.org/wiki/File:Seal_of_Bihar.svg).

```{r tigertest, include=FALSE}
tiger <- image_read_svg('http://jeroen.github.io/images/tiger.svg', width = 350)
print(tiger)
```

```{r upseal, echo=FALSE}
bihar_seal <- image_read_svg('https://upload.wikimedia.org/wikipedia/commons/e/e9/Seal_of_Bihar.svg', width=478)
bihar_seal
```

```{r, include=FALSE}
# Seal svg rendered into png format. 
image_write(bihar_seal, path = "bihar_seal.png", format = "png")
bihar_seal_png <- image_convert(bihar_seal, "png")
image_info(bihar_seal_png)
bihar_seal_png
```

```{r, include=FALSE}
# Svg rendered into bitmap form. 
# Or: convert into raw bitmap data
bitmap <- rsvg_raw('https://upload.wikimedia.org/wikipedia/commons/e/e9/Seal_of_Bihar.svg', width = 478)
str(bitmap)
# Read the bitmap in magick
biharbitmap <- magick::image_read(bitmap)
biharbitmap
```

[Wikipedia](https://en.wikipedia.org/wiki/Emblem_of_Bihar#:~:text=The%20emblem%20of%20Bihar%20depicts,%22Bihar%22%20inscribed%20on%20it.) suggests the seal:

> depicts the Bodhi tree with prayer beads, flanked by two swastikas. The base of the tree is a brick with the letter Bihar inscribed on it that means bihar.

## 2. State district map 

* Original file size: 512x380 pixels and 269KB. Image source: [Nirinsanity](https://commons.wikimedia.org/wiki/File:Bihar_districts.svg).

```{r district, echo=FALSE}
district <- image_read_svg('https://upload.wikimedia.org/wikipedia/commons/a/a9/Bihar_districts.svg', height=478)
district
```

```{r districtpng, include=FALSE}
image_write(district, path = "district.png", format = "png")
district_png <- image_convert(district, "png")
```

```{r distpng, include=FALSE}
district_png
```

[Wikipedia](https://en.wikipedia.org/wiki/List_of_districts_of_Bihar) details 38 districts. 

```{r rawdatacollect, include=FALSE}
# get the URL for the wikipedia page with all states
url <- "https://en.wikipedia.org/wiki/List_of_districts_of_Bihar"

# use that URL to scrape the states table using rvest
districttable <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[2]') %>% 
  html_table(fill=TRUE)
#create a vector of states
head(districttable)

```

```{r convertchar2num, include=FALSE}
# convert character column to numeric in three steps
# step 1 is population column - remove
# step 2 is convert to numeric
# step 3 is convert other char columns to numeric directly
distdf <- as.data.frame(districttable)
# remove unwanted columns
distdf <- distdf[,-c(1,8)]
colnames(distdf)
# tidy column names with janitor
disttidy <- distdf %>% 
  janitor::clean_names()
disttidy
# convert character columns to numeric
disttidy[,"population_2011"] <- gsub(",", "", disttidy[,"population_2011"])
disttidy[,"area_km"] <- gsub(",", "", disttidy[,"area_km"])
disttidy[,"density_km_2011"] <- gsub(",", "", disttidy[,"density_km_2011"])
disttidy[,"population_2011"] <- as.numeric(disttidy[,"population_2011"])
disttidy[,"area_km"] <- as.numeric(disttidy[,"area_km"])
disttidy[,"density_km_2011"] <- as.numeric(disttidy[,"density_km_2011"])
disttidy

# order columns
disttidy2 <- disttidy[,c(2:6,1)]
head(disttidy2)
```

```{r, include=FALSE}
distorderpop <- disttidy2[
  order(-disttidy2[,3]),
  ]
distorderpop

```

\newpage

```{r distorderpop, echo=FALSE}
knitr::kable(distorderpop[,c(1:6)], "html",
             caption = "Bihar districts by 2011 population",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

```{r, include=FALSE}
distordername <- disttidy2[
  order(disttidy2[,1]),
  ]
distordername

```

\newpage

```{r distordername, echo=FALSE}
knitr::kable(distordername[,c(1:6)], "html",
             caption = "Bihar districts by name",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

## 3. State legislature

### Upper house 

* The state legislature is bicameral. The upper house of 75 members is called the legislative council or Vidhan Parishad^[https://en.wikipedia.org/wiki/Bihar_Legislative_Council]. 

* Term for Vidhan Parishad members is 6 years. Seats are drawn from five groups: legislative assembly, local authority, nominated members, graduate and teachers. The distribution of seats is shown in table. 


```{r rs_seats, include=FALSE}
rs_seats <- data.frame("Constituency"=c("Graduate", 
                                        "Teachers", 
                                        "Local Authority",
                                        "Legislative assembly",
                                        "Nominated members"), 
                     "Seats" = c(6, 6, 24, 27, 12))

rs_seats
```

```{r, include=FALSE}
rsseatorder <- rs_seats[
  order(-rs_seats[,2]),
  ]
```

```{r, include=FALSE}
rsseatorder
```

```{r upperhousetable, echo=FALSE}
knitr::kable(rsseatorder, "html",
             caption = "Bihar upper house by constituency",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

### Lower house 

* The lower house has 243 members and is called the legislative assembly or Vidhan Sabha^[https://en.wikipedia.org/wiki/Bihar_Legislative_Assembly].

* Term for Vidhan Sabha members is 5 years. Latest election was 2017. Upcoming election scheduled for 2022. 

* Janata Dal (United) party has 73 seats and BJP has 53 seats. Rashtriya Janata Dal and Indian National Congress have 80 and 26 seats. 

* Vidhan Sabha constituencies and elected members as of 2015^[https://en.wikipedia.org/wiki/2015_Bihar_Legislative_Assembly_election]

* Original file size: 953x1280 pixels and 8850KB. Image source: [Furfur](https://fr.m.wikipedia.org/wiki/Fichier:Wahlkreise_zur_Vidhan_Sabha_von_Bihar.svg).

```{r constituency, echo=FALSE}
constituency <- image_read_svg('https://upload.wikimedia.org/wikipedia/commons/9/9b/Wahlkreise_zur_Vidhan_Sabha_von_Bihar.svg', height=478)
constituency
```

```{r constituencypng, include=FALSE}
image_write(constituency, path = "constituency.png", format = "png")
constituency_png <- image_convert(constituency, "png")
```

```{r constitpng, include=FALSE}
constituency_png
```

## 5. Electoral constituencies

```{r biharlowerresults, include=FALSE}
# get the URL for the wikipedia page with all states
vsurl <- "https://en.wikipedia.org/wiki/2015_Bihar_Legislative_Assembly_election"

# use that URL to scrape the states table using rvest
vselection <- vsurl %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[10]') %>% 
  html_table(fill=TRUE)
#create a vector of election results
head(vselection)

```

```{r, include=FALSE}
vselection_df <- as.data.frame(vselection)
vsdf <- vselection_df[,c(1,3,2,6)]
# tidy column names with janitor
vstidy <- vsdf %>% 
  janitor::clean_names()
vstidy

colnames(vstidy)[1] <- "id.no"
colnames(vstidy)[3] <- "constituency"
vstidy
```

```{r, include=FALSE}
vsdforder <- vstidy[
  order(vstidy[,2]),
  ]
```

```{r, include=FALSE}
vsdforder
```

```{r vstable, echo=FALSE}
knitr::kable(vsdforder, "html",
             caption = "Bihar lower house by constituency",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

## 6. Constituencies by district  

```{r, include=FALSE}
d <- distinct(disttidy2[1])
```

```{r, include=FALSE}
c <-distinct(vstidy[2])
```

```{r, include=FALSE}
identical(d, c)
```

```{r, include=FALSE}
# identify differences between the two data table district columns
dplyr::anti_join(d, c, by = "district")
```

```{r, include=FALSE}
# edit cells identified as different so as to obtain consistency between data sets 
# change Bagpat in distup dataframe to Baghpat

disttidy2 <- disttidy2 %>%
     mutate(district=replace(district, district == "East Champaran", "Purvi Champaran")) %>%
     as.data.frame()

disttidy2 <- disttidy2 %>%
     mutate(district=replace(district, district == "West Champaran", "Paschim Champaran")) %>%
     as.data.frame()

```

```{r, include=FALSE}
# make a join of ml and mlc
dc <- dplyr::full_join(disttidy2, vstidy, by = "district")
dc
```

```{r, include=FALSE}
#select columns required
dc2 <- dc[,c(1,3,4,5,8)]

```

```{r, include=FALSE}
# merge count no of constituencies by district
# make district and division factors
dc2$district <- as.factor(dc2$district)
head(dc2)
```

```{r, include=FALSE}
# count constituencies per district
table(dc2$district)
prop.table(table(dc2$district))

```

```{r, include=FALSE}
# alternate method using dplyr to generate a frequency table of constituencies by district
bihar_constit <- dc2 %>% count(district, sort = TRUE)
```

```{r, include=FALSE}
bihar_constit
```

```{r, include=FALSE}
# bind constituency count to upd data frame 
dc2merge <- merge(disttidy2, bihar_constit, by.disttidy2 = "district", by.bihar_constit = "district", all = TRUE)


# rename constituency count column
colnames(dc2merge)[7] <- "constituency_count"
dc2merge

```

```{r dupconcount, echo=FALSE}
knitr::kable(dc2merge[,c(1:5,7)], "html",
             caption = "Bihar constituency count by district",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

## 7. India parliamentary representation 

* Wikipedia^[https://en.wikipedia.org/wiki/Parliament_of_India] describes India's legislative body as:

> bicameral with 788 seats divided as to Rajya Sabha (Council of States) of 245 members, and, the Lok Sabha (House of the People) of 543 members all presided over by President of India;

> Lok Sabha members directly elected by public who vote in single member districts.

> Rajya Sabha members elected by State Legislative assembly by proportional representation. 

* For **Bihar** there are 40 Lok Sabha seats and 16 Rajya Sabha seats representing 9% of total parliamentary seats. 

* Original file size: 3818x4600 pixels and 75KB. PNG creator: [The,Manish](https://commons.wikimedia.org/wiki/File:Indian_Parliament_Seat_Composition.png).

```{r indparl, echo=FALSE}
# Example image

indparl <- image_read("https://upload.wikimedia.org/wikipedia/commons/6/6e/Indian_Parliament_Seat_Composition.png")
image_scale(indparl, "x478") # height: 300px
```


## 8. Lok Sabha election past

* [Wikipedia](https://en.wikipedia.org/wiki/Elections_in_India#History_of_Lok_Sabha_Elections) details 17 Lok Sabha elections since 1951 with the latest being that of 2019. 

* The most recent election result of 2019 gave 303 seats to BJP. 

* Next Lok Sabha elections scheduled for 2024. 

```{r lsresults, include=FALSE}
# get the URL for the wikipedia page with all states
url2 <- "https://en.wikipedia.org/wiki/Elections_in_India#History_of_Lok_Sabha_Elections"

# use that URL to scrape the states table using rvest
lselections <- url2 %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[3]') %>% 
  html_table(fill=TRUE)
#create a vector of election results
head(lselections)

```

```{r, include=FALSE}
lok_df <- as.data.frame(lselections)

```

\newpage

```{r loksabha, echo=FALSE}
knitr::kable(lok_df[,c(1:5)], "html",
             caption = "Lok Sabha election results",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

## 9. Rajya Sabha composition by party 

* After the past week's election the BJP party and its alliance National Democratic Alliance (NDA) have majority in Rajya Sabha with 84 seats and 114 seats. Congress have 41 seats and its alliance United Progressive Alliance has 62 seats in total^[Source: https://en.wikipedia.org/wiki/Rajya_Sabha#Membership_by_party].

* Elections occur as and when members retire and each term is for 6 years. Next elections scheduled for 2021. 

```{r rsresults, include=FALSE}
# get the URL for the wikipedia page with all states
url3 <- "https://en.wikipedia.org/wiki/Rajya_Sabha#Membership_by_party"

# use that URL to scrape the states table using rvest
rselections <- url3 %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[3]') %>% 
  html_table(fill=TRUE)
#create a vector of election results
head(rselections)

```

```{r, include=FALSE}
rajya_df <- as.data.frame(rselections)

```

\newpage

```{r rajyasabha, echo=FALSE}
knitr::kable(rajya_df[,c(1:3)], "html",
             caption = "2020-06-19 Rajya Sabha composition by party",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```

