---
title: India district broad data
author: michael
date: '2020-06-19'
slug: india-district-broad-data
categories:
  - R
tags:
  - India
  - Population
  - Geography
  - Census
---

### Task:

1. Extract table data of India districts from [Wikipedia](https://en.wikipedia.org/wiki/List_of_districts_in_India#:~:text=In%20some%20cases%20districts%20are,the%202001%20Census%20of%20India)
1. Combine each table rows into one dataframe;
1. add a column to combined data frame that provides name of original dataframe - so as to facilitate identifying source of data
1. reorder the columns in expanded data frame to put original data frame column first. 

There are 36 tables of India states and territories listed on the wikipedia page. Each table was extracted and stored in a separate data frame. A comma separated list of individual data frames was put into a vector. And a function then appended a column with the name of individual source data frame to each row of data. After that the individual data frames were bounded into one object and variable columns were reordered. 

Difficulty and cumbersome script used to allow for unwanted notes included within population, area, and, density cells. More efficient method is required to remove the commas from character variable to permit easy conversion to numeric form. 

The value add in result is to aggregate the total population, area and compute population density by state and territory. And then to provide a complete listing of all data by district for each state. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
pkgs = c("rvest", "tidyverse", "janitor", "lubridate", "dplyr", "kableExtra", "data.table", "forcats", "car", "plyr", "magrittr") 
# lubridate for %m+% function
# rvest for web scraping
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

```{r datevars, include=FALSE}
# save current system date to a variable
today <- Sys.Date()
# subtract 3 months from the current date
date <- today %m+% months(-3)
print(date)
```

```{r rawdatacollect, include=FALSE}
# get the URL for the wikipedia page with all states
url <- "https://en.wikipedia.org/wiki/List_of_districts_in_India#:~:text=In%20some%20cases%20districts%20are,the%202001%20Census%20of%20India."

# use that URL to scrape the states table using rvest
andhrapradesh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[3]') %>% 
  html_table(fill=TRUE)
#create a vector of states
andhrapradesh <- andhrapradesh.m[[1]]

arunachalpradesh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[4]') %>% 
  html_table(fill=TRUE)
#create a vector of states
arunachalpradesh <- arunachalpradesh.m[[1]]

assam.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[5]') %>% 
  html_table(fill=TRUE)
#create a vector of states
assam <- assam.m[[1]]

bihar.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[6]') %>% 
  html_table(fill=TRUE)
#create a vector of states
bihar <- bihar.m[[1]]

chhattisgarh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[7]') %>% 
  html_table(fill=TRUE)
#create a vector of states
chhattisgarh <- chhattisgarh.m[[1]]

goa.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[8]') %>% 
  html_table(fill=TRUE)
#create a vector of states
goa <- goa.m[[1]]

gujarat.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[9]') %>% 
  html_table(fill=TRUE)
#create a vector of states
gujarat <- gujarat.m[[1]]

haryana.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[10]') %>% 
  html_table(fill=TRUE)
#create a vector of states
haryana <- haryana.m[[1]]

himachalpradesh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[11]') %>% 
  html_table(fill=TRUE)
#create a vector of states
himachalpradesh <- himachalpradesh.m[[1]]

jharkhand.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[12]') %>% 
  html_table(fill=TRUE)
#create a vector of states
jharkhand <- jharkhand.m[[1]]

karnataka.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[13]') %>% 
  html_table(fill=TRUE)
#create a vector of states
karnataka <- karnataka.m[[1]]

kerala.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[14]') %>% 
  html_table(fill=TRUE)
#create a vector of states
kerala <- kerala.m[[1]]

madhyapradesh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[15]') %>% 
  html_table(fill=TRUE)
#create a vector of states
madhyapradesh <- madhyapradesh.m[[1]]

maharashtra.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[16]') %>% 
  html_table(fill=TRUE)
#create a vector of states
maharashtra <- maharashtra.m[[1]]

manipur.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[17]') %>% 
  html_table(fill=TRUE)
#create a vector of states
manipur <- manipur.m[[1]]

meghalaya.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[18]') %>% 
  html_table(fill=TRUE)
#create a vector of states
meghalaya <- meghalaya.m[[1]]

mizoram.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[19]') %>% 
  html_table(fill=TRUE)
#create a vector of states
mizoram <- mizoram.m[[1]]

nagaland.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[20]') %>% 
  html_table(fill=TRUE)
#create a vector of states
nagaland <- nagaland.m[[1]]

odisha.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[21]') %>% 
  html_table(fill=TRUE)
#create a vector of states
odisha <- odisha.m[[1]]

punjab.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[22]') %>% 
  html_table(fill=TRUE)
#create a vector of states
punjab <- punjab.m[[1]]

rajasthan.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[23]') %>% 
  html_table(fill=TRUE)
#create a vector of states
rajasthan <- rajasthan.m[[1]]

sikkim.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[24]') %>% 
  html_table(fill=TRUE)
#create a vector of states
sikkim <- sikkim.m[[1]]

tamilnadu.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[25]') %>% 
  html_table(fill=TRUE)
#create a vector of states
tamilnadu <- tamilnadu.m[[1]]

telangana.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[26]') %>% 
  html_table(fill=TRUE)
#create a vector of states
telangana <- telangana.m[[1]]

tripura.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[27]') %>% 
  html_table(fill=TRUE)
#create a vector of states
tripura <- tripura.m[[1]]

uttarpradesh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[28]') %>% 
  html_table(fill=TRUE)
#create a vector of states
uttarpradesh <- uttarpradesh.m[[1]]

uttarakhand.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[29]') %>% 
  html_table(fill=TRUE)
#create a vector of states
uttarakhand <- uttarakhand.m[[1]]

westbengal.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[30]') %>% 
  html_table(fill=TRUE)
#create a vector of states
westbengal <- westbengal.m[[1]]

andamannicobar.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[31]') %>% 
  html_table(fill=TRUE)
#create a vector of states
andamannicobar <- andamannicobar.m[[1]]

chandigarh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[32]') %>% 
  html_table(fill=TRUE)
#create a vector of states
chandigarh <- chandigarh.m[[1]]

dadranagardaman.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[33]') %>% 
  html_table(fill=TRUE)
#create a vector of states
dadranagardaman <- dadranagardaman.m[[1]]

jammukashmir.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[34]') %>% 
  html_table(fill=TRUE)
#create a vector of states
jammukashmir <- jammukashmir.m[[1]]

ladakh.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[35]') %>% 
  html_table(fill=TRUE)
#create a vector of states
ladakh <- ladakh.m[[1]]

lakshadweep.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[36]') %>% 
  html_table(fill=TRUE)
#create a vector of states
lakshadweep <- lakshadweep.m[[1]]

nct.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[37]') %>% 
  html_table(fill=TRUE)
#create a vector of states
nct <- nct.m[[1]]

puducherry.m <- url %>%
# read the HTML from the webpage
  read_html() %>%
  # one way to get table
  #html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]') %>%
  # easier way to get table
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[4]/div/table[38]') %>% 
  html_table(fill=TRUE)
#create a vector of states
puducherry <- puducherry.m[[1]]

```

```{r, include=FALSE}
# add column for variable name in each df

ID <- c(
"andhrapradesh",
"arunachalpradesh",
"assam",
"bihar",
"chhattisgarh",
"goa",
"gujarat",
"haryana",
"himachalpradesh",
"jharkhand",
"karnataka",
"kerala",
"madhyapradesh",
"maharashtra",
"manipur",
"meghalaya",
"mizoram",
"nagaland",
"odisha",
"punjab",
"rajasthan",
"sikkim",
"tamilnadu",
"telangana",
"tripura",
"uttarpradesh",
"uttarakhand",
"westbengal",
"andamannicobar",
"chandigarh",
"dadranagardaman",
"jammukashmir",
"ladakh",
"lakshadweep",
"nct",
"puducherry")

```

```{r colheadconsistent, include=FALSE}
#make column headers consistent
dfList <- list(c(andhrapradesh,
                 arunachalpradesh,
                 assam,
                 bihar,
                 chhattisgarh,
                 goa,
                 gujarat,
                 haryana,
                 himachalpradesh,
                 jharkhand,
                 karnataka,
                 kerala,
                 madhyapradesh,
                 maharashtra,
                 manipur,
                 meghalaya,
                 mizoram,
                 nagaland,
                 odisha,
                 punjab,
                 rajasthan,
                 sikkim,
                 tamilnadu,
                 telangana,
                 tripura,
                 uttarpradesh,
                 uttarakhand,
                 westbengal,
                 andamannicobar,
                 chandigarh,
                 dadranagardaman,
                 jammukashmir,
                 ladakh,
                 lakshadweep,
                 nct,
                 puducherry)
               )

colnames <- c("#", "Code", "District", "Headquarters", "Pop_2011", "Area_km2", "Density_/km2") 

dfList <- list(
  andhrapradesh = andhrapradesh,
  arunachalpradesh = arunachalpradesh,
  assam = assam,
  bihar = bihar,
  chhattisgarh = chhattisgarh,
  goa = goa,
  gujarat = gujarat,
  haryana = haryana,
  himachalpradesh = himachalpradesh,
  jharkhand = jharkhand,
  karnataka = karnataka,
  kerala = kerala,
  madhyapradesh = madhyapradesh,
  maharashtra = maharashtra,
  manipur = manipur,
  meghalaya = meghalaya,
  mizoram = mizoram,
  nagaland = nagaland,
  odisha = odisha,
  punjab = punjab,
  rajasthan = rajasthan,
  sikkim = sikkim,
  tamilnadu = tamilnadu,
  telangana = telangana,
  tripura = tripura,
  uttarpradesh = uttarpradesh,
  uttarakhand = uttarakhand,
  westbengal = westbengal,
  andamannicobar = andamannicobar,
  chandigarh = chandigarh,
  dadranagardaman = dadranagardaman,
  jammukashmir = jammukashmir,
  ladakh = ladakh,
  lakshadweep = lakshadweep,
  nct = nct,
  puducherry = puducherry)

list2env(lapply(dfList, setNames, colnames), .GlobalEnv)
```

```{r convertchar2num, include=FALSE}
# convert character column to numeric in three steps
# step 1 is population column - remove commans
# step 2 is convert to numeric
# step 3 is convert other char columns to numeric directly

andhrapradesh[,"Pop_2011"] <- gsub(",", "", andhrapradesh[,"Pop_2011"])
andhrapradesh[,"Pop_2011"] <- as.numeric(andhrapradesh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
andhrapradesh[,coltype] <- 
                     lapply(andhrapradesh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

arunachalpradesh[,"Pop_2011"] <- gsub(",", "", arunachalpradesh[,"Pop_2011"])
arunachalpradesh[,"Pop_2011"] <- as.numeric(arunachalpradesh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
arunachalpradesh[,coltype] <- 
                     lapply(arunachalpradesh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

assam[,"Pop_2011"] <- gsub(",", "", assam[,"Pop_2011"])
assam[,"Pop_2011"] <- as.numeric(assam[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
assam[,coltype] <- 
                     lapply(assam[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

bihar[,"Pop_2011"] <- gsub(",", "", bihar[,"Pop_2011"])
bihar[,"Pop_2011"] <- as.numeric(bihar[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
bihar[,coltype] <- 
                     lapply(bihar[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

chandigarh[,"Pop_2011"] <- gsub(",", "", chandigarh[,"Pop_2011"])
chandigarh[,"Pop_2011"] <- as.numeric(chandigarh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
chandigarh[,coltype] <- 
                     lapply(chandigarh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

andamannicobar[,"Pop_2011"] <- gsub(",", "", andamannicobar[,"Pop_2011"])
andamannicobar[,"Pop_2011"] <- as.numeric(andamannicobar[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
andamannicobar[,coltype] <- 
                     lapply(andamannicobar[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

chhattisgarh[,"Pop_2011"] <- gsub(",", "", chhattisgarh[,"Pop_2011"])
chhattisgarh[,"Pop_2011"] <- as.numeric(chhattisgarh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
chhattisgarh[,coltype] <- 
                     lapply(chhattisgarh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

dadranagardaman[,"Pop_2011"] <- gsub(",", "", dadranagardaman[,"Pop_2011"])
dadranagardaman[,"Pop_2011"] <- as.numeric(dadranagardaman[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
dadranagardaman[,coltype] <- 
                     lapply(dadranagardaman[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

goa[,"Pop_2011"] <- gsub(",", "", goa[,"Pop_2011"])
goa[,"Pop_2011"] <- as.numeric(goa[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
goa[,coltype] <- 
                     lapply(goa[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

gujarat[,"Pop_2011"] <- gsub(",", "", gujarat[,"Pop_2011"])
gujarat[,"Pop_2011"] <- as.numeric(gujarat[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
gujarat[,coltype] <- 
                     lapply(gujarat[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

haryana[,"Pop_2011"] <- gsub(",", "", haryana[,"Pop_2011"])
haryana[,"Pop_2011"] <- as.numeric(haryana[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
haryana[,coltype] <- 
                     lapply(haryana[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

himachalpradesh[,"Pop_2011"] <- gsub(",", "", himachalpradesh[,"Pop_2011"])
himachalpradesh[,"Pop_2011"] <- as.numeric(himachalpradesh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
himachalpradesh[,coltype] <- 
                     lapply(himachalpradesh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

jammukashmir[,"Pop_2011"] <- gsub(",", "", jammukashmir[,"Pop_2011"])
jammukashmir[,"Pop_2011"] <- as.numeric(jammukashmir[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
jammukashmir[,coltype] <- 
                     lapply(jammukashmir[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

jharkhand[,"Pop_2011"] <- gsub(",", "", jharkhand[,"Pop_2011"])
jharkhand[,"Pop_2011"] <- as.numeric(jharkhand[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
jharkhand[,coltype] <- 
                     lapply(jharkhand[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

karnataka[,"Pop_2011"] <- gsub(",", "", karnataka[,"Pop_2011"])
karnataka[,"Pop_2011"] <- as.numeric(karnataka[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
karnataka[,coltype] <- 
                     lapply(karnataka[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

kerala[,"Pop_2011"] <- gsub(",", "", kerala[,"Pop_2011"])
kerala[,"Pop_2011"] <- as.numeric(kerala[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
kerala[,coltype] <- 
                     lapply(kerala[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

ladakh[,"Pop_2011"] <- gsub(",", "", ladakh[,"Pop_2011"])
ladakh[,"Pop_2011"] <- as.numeric(ladakh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
ladakh[,coltype] <- 
                     lapply(ladakh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

lakshadweep[,"Pop_2011"] <- gsub(",", "", lakshadweep[,"Pop_2011"])
lakshadweep[,"Pop_2011"] <- as.numeric(lakshadweep[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
lakshadweep[,coltype] <- 
                     lapply(lakshadweep[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

madhyapradesh[,"Pop_2011"] <- gsub(",", "", madhyapradesh[,"Pop_2011"])
madhyapradesh[,"Pop_2011"] <- as.numeric(madhyapradesh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
madhyapradesh[,coltype] <- 
                     lapply(madhyapradesh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

maharashtra[,"Pop_2011"] <- gsub(",", "", maharashtra[,"Pop_2011"])
maharashtra[,"Pop_2011"] <- as.numeric(maharashtra[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
maharashtra[,coltype] <- 
                     lapply(maharashtra[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

manipur[,"Pop_2011"] <- gsub(",", "", manipur[,"Pop_2011"])
manipur[,"Pop_2011"] <- as.numeric(manipur[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
manipur[,coltype] <- 
                     lapply(manipur[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

meghalaya[,"Pop_2011"] <- gsub(",", "", meghalaya[,"Pop_2011"])
meghalaya[,"Pop_2011"] <- as.numeric(meghalaya[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
meghalaya[,coltype] <- 
                     lapply(meghalaya[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

mizoram[,"Pop_2011"] <- gsub(",", "", mizoram[,"Pop_2011"])
mizoram[,"Pop_2011"] <- as.numeric(mizoram[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
mizoram[,coltype] <- 
                     lapply(mizoram[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

nagaland[,"Pop_2011"] <- gsub(",", "", nagaland[,"Pop_2011"])
nagaland[,"Pop_2011"] <- as.numeric(nagaland[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
nagaland[,coltype] <- 
                     lapply(nagaland[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

nct[,"Pop_2011"] <- gsub(",", "", nct[,"Pop_2011"])
nct[,"Pop_2011"] <- as.numeric(nct[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
nct[,coltype] <- 
                     lapply(nct[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

odisha[,"Pop_2011"] <- gsub(",", "", odisha[,"Pop_2011"])
odisha[,"Pop_2011"] <- as.numeric(odisha[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
odisha[,coltype] <- 
                     lapply(odisha[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

puducherry[,"Pop_2011"] <- gsub(",", "", puducherry[,"Pop_2011"])
puducherry[,"Pop_2011"] <- as.numeric(puducherry[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
puducherry[,coltype] <- 
                     lapply(puducherry[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

punjab[,"Pop_2011"] <- gsub(",", "", punjab[,"Pop_2011"])
punjab[,"Pop_2011"] <- as.numeric(punjab[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
punjab[,coltype] <- 
                     lapply(punjab[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

rajasthan[,"Pop_2011"] <- gsub(",", "", rajasthan[,"Pop_2011"])
rajasthan[,"Pop_2011"] <- as.numeric(rajasthan[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
rajasthan[,coltype] <- 
                     lapply(rajasthan[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

sikkim[,"Pop_2011"] <- gsub(",", "", sikkim[,"Pop_2011"])
sikkim[,"Pop_2011"] <- as.numeric(sikkim[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
sikkim[,coltype] <- 
                     lapply(sikkim[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

tamilnadu[,"Pop_2011"] <- gsub(",", "", tamilnadu[,"Pop_2011"])
tamilnadu[,"Pop_2011"] <- as.numeric(tamilnadu[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
tamilnadu[,coltype] <- 
                     lapply(tamilnadu[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

telangana[,"Pop_2011"] <- gsub(",", "", telangana[,"Pop_2011"])
telangana[,"Pop_2011"] <- as.numeric(telangana[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
telangana[,coltype] <- 
                     lapply(telangana[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

tripura[,"Pop_2011"] <- gsub(",", "", tripura[,"Pop_2011"])
tripura[,"Pop_2011"] <- as.numeric(tripura[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
tripura[,coltype] <- 
                     lapply(tripura[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

uttarakhand[,"Pop_2011"] <- gsub(",", "", uttarakhand[,"Pop_2011"])
uttarakhand[,"Pop_2011"] <- as.numeric(uttarakhand[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
uttarakhand[,coltype] <- 
                     lapply(uttarakhand[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

uttarpradesh[,"Pop_2011"] <- gsub(",", "", uttarpradesh[,"Pop_2011"])
uttarpradesh[,"Pop_2011"] <- as.numeric(uttarpradesh[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
uttarpradesh[,coltype] <- 
                     lapply(uttarpradesh[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))

westbengal[,"Pop_2011"] <- gsub(",", "", westbengal[,"Pop_2011"])
westbengal[,"Pop_2011"] <- as.numeric(westbengal[,"Pop_2011"])
coltype <- c("Area_km2", "Density_/km2")
westbengal[,coltype] <- 
                     lapply(westbengal[,coltype], function(i) 
                                                          as.numeric(sub(',', '', i)))
```

```{r addbackmissdata, include=FALSE}
# add back cells where data deleted in conversion to numeric
str(lakshadweep)
head(lakshadweep)
lakshadweep[,c(5:7)]=c(64473, 30, 2149)
head(lakshadweep)

# add back cells where data deleted in conversion to numeric
head(dadranagardaman)
dadranagardaman[c(1:2),c(5:6)]=c(191173, 52074, 72, 39)
head(dadranagardaman)

# add back cells where data deleted in conversion to numeric
head(puducherry)
puducherry[c(1:4),c(5:7)]=c(200222, 41816, 950289, 55626, 157, 9, 293, 30, 1275, 4646, 3232, 1854)
head(puducherry)

# add back cells where data deleted in conversion to numeric
head(assam)
assam[3,5]=c(612491)
assam[16,5]=c(931218)
assam[24,5]=c(167304)
assam[29,5]=c(555114)
assam[33,5]=c(300320)
head(assam)

```

```{r makecolwithobjectname, include=FALSE}
#make new column with name of object as a column with name in every row

bind_rows(lst(andhrapradesh,
                 arunachalpradesh,
                 assam,
                 bihar,
                 chhattisgarh,
                 goa,
                 gujarat,
                 haryana,
                 himachalpradesh,
                 jharkhand,
                 karnataka,
                 kerala,
                 madhyapradesh,
                 maharashtra,
                 manipur,
                 meghalaya,
                 mizoram,
                 nagaland,
                 odisha,
                 punjab,
                 rajasthan,
                 sikkim,
                 tamilnadu,
                 telangana,
                 tripura,
                 uttarpradesh,
                 uttarakhand,
                 westbengal,
                 andamannicobar,
                 chandigarh,
                 dadranagardaman,
                 jammukashmir,
                 ladakh,
                 lakshadweep,
                 nct,
                 puducherry),.id="state_terr")
```

```{r appendfunction, include=FALSE}

# Function to combine (rbind) data frames and create column with name of original data frames

# Source: https://stackoverflow.com/questions/15162197/combine-rbind-data-frames-and-create-column-with-name-of-original-data-frames

AppendMe <- function(dfNames) {
  do.call(rbind, lapply(dfNames, function(x) {
    cbind(get(x), source = x)
  }))
}
```

```{r, include=FALSE }
# append column to dataset with name of source variable in rbind operations
bodywhole2 <- AppendMe(c(ID))
```


```{r, include=FALSE}
# get column names
# source: http://www.sthda.com/english/wiki/reordering-data-frame-columns-in-r
colnames(bodywhole2)

```

```{r, include=FALSE}
bodywhole3 <- bodywhole2[, c(8, 1, 2, 3,4,5,6,7)]
tail(bodywhole3)
```

```{r arrangedtable, include=FALSE}
# source: https://chartio.com/resources/tutorials/how-to-sort-a-data-frame-by-multiple-columns-in-r/

# first sort by source and then by state+terr

tab <- bodywhole3 [
  order( bodywhole3[,1], desc(bodywhole3[,6])),
]

```

```{r fs, include=FALSE}
fwrite(tab, file = "/Users/michaelleary/Documents/GitHub/sources/tab.csv", row.names = TRUE) # in csv form

```

```{r char2factor, include=FALSE}
# generate a summary table by state and terr
# generate a list of character variables within data frame
tab %>%
sapply(is.character) %>%
which() %T>%
print() ->
chari
str(chari)

# make a vector of character variables

tab %>%
names() %>%
'['(chari) %T>%
print() ->
charc
str(charc)

# convert the source character variable to factor
tab[charc[1]] %<>% map(factor)  #uncheck
str(tab)

```

```{r districtcount, include=FALSE}
tab[1] %>%
  group_by(source) %>%
  count(.) 

```

```{r sumtable, include=FALSE}
str(tab)
sumtable <- tab %>%
  group_by(source) %>%
  dplyr::summarize(total_pop = sum(Pop_2011, na.rm=TRUE),
                   total_area = round(sum(Area_km2, na.rm=TRUE))) %>%
  mutate(pop_per_km2 = round(total_pop / total_area)) %>%
  arrange(desc(total_pop))
  

```


```{r sumtableprint, echo=FALSE}
knitr::kable(sumtable, "html",
             caption = "India state and territory: pop, area",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```


\newpage

```{r tabtable, echo=FALSE}
knitr::kable(tab[,c(1,4,6,8)], "html",
             caption = "India district: pop, area",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```
