---
title: Food consumption by country
author: michael
date: '2020-06-06'
slug: food-consumption-by-country
categories:
  - R
tags:
  - Agriculture
  - Food
---

Output from this script provides charts of consumption of various food categories (in terms of kg/person/year) by country. The data is sourced from Food and Agriculture Organization of the United Nations as of 2018 and  published by [nu3](https://www.nu3.de/blogs/nutrition/food-carbon-footprint-index-2018). 

The script extracts an html table data from the website, tidies the extracted data, and, generates charts. The script is virtually all written by [Kasia Kulma](https://r-tastic.co.uk/post/from-messy-to-tidy/).  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
pkgs = c("rvest", 
         "dplyr", 
         "tidyr",
         "stringr",
         "janitor", 
         "ggplot2",
         "forcats",
         "magrittr",
         "purrr",
         "knitr", 
         "kableExtra")
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

```{r extractwebsitedata}
url <- "https://www.nu3.de/blogs/nutrition/food-carbon-footprint-index-2018"

```

```{r websitehtmlscrape, include=FALSE}
url_html <- read_html(url)

```

```{r extracthtmltable, include=FALSE}
whole_table <- url_html %>% 
 html_nodes('table') %>%
 html_table(fill = TRUE) %>%
 .[[1]]

```

```{r exmainehtmltabledata, include=FALSE}
str(whole_table)
```

```{r tidyhtmltableheaders, include=FALSE}

# using rvest package
table_content <- whole_table %>%
 select(-X1) %>% # remove redundant column
 filter(!dplyr::row_number() %in% 1:3) # remove redundant rows

head(table_content)
```

```{r extracttitleattributes, include=FALSE}
raw_headers <- url_html %>%
 html_nodes(".thead-icon") %>%
 html_attr('title')

raw_headers[1:35]
```

```{r selectbottheaders, include=FALSE}
tidy_bottom_header <- raw_headers[28:length(raw_headers)]
tidy_bottom_header[1:10]
```

```{r selectmidheaders, include=FALSE}
raw_middle_header <- raw_headers[17:27]
raw_middle_header
```

```{r sortcolnamestovarorder, include=FALSE}
# organize column names to reflect variable order
tidy_headers <- c(
 rep(raw_middle_header[1:7], each = 2),
 "animal_total",
 rep(raw_middle_header[8:length(raw_middle_header)], each = 2),
 "non_animal_total",
 "country_total")

tidy_headers

```

```{r combinecolname, include=FALSE}
combined_colnames <- paste(tidy_headers, tidy_bottom_header, sep = ';')
colnames(table_content) <- c("Country", combined_colnames)
glimpse(table_content[, 1:10])
```

```{r, include=FALSE}
head(table_content)
```

```{r, include=FALSE}
long_table <- table_content %>%
 # make column names observations of Category variable
 tidyr::pivot_longer(cols = -Country, names_to = "Category", values_to = "Values") %>%
 # separate food-related information from the metric
 tidyr::separate(col = Category, into = c("Food Category", "Metric"), sep = ';')

glimpse(long_table)
```

```{r, include=FALSE}
head(long_table)
```

```{r tidycolname, include=FALSE}
tidy_table <- long_table %>%
   tidyr::pivot_wider(names_from = Metric, values_from = Values) %>%
   janitor::clean_names('snake')

glimpse(tidy_table)
```

```{r finaltable, include=FALSE}
final_table <- tidy_table %>%
 rename(consumption = 3,
         co2_emmission = 4) %>%
 filter(!stringr::str_detect(food_category, "total"))

head(final_table, 20)
```

```{r makefoodfactor, include=FALSE}
final_table2 <- final_table
char <- c("food_category")
final_table2[char] %<>% map(factor)
glimpse(final_table2)
```

```{r makenumericvar, include=FALSE}
#identify numeric variables to process
cvars <- c("consumption", "co2_emmission")
#check current class of variables
final_table2[cvars] %>% sapply(class)
#To convert to numeric
final_table2[cvars] %>% sapply(as.numeric) #uncheck when needed
# convert to numeric - then check and remote NA
final_table2[cvars] %<>% map(as.numeric)
glimpse(final_table2)
```

```{r usplot, include=FALSE}
us_foodplot <- final_table2 %>% 
  filter(country %in% c("USA")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
us_foodplot
```

```{r chinaplot, include=FALSE}
china_foodplot <- final_table2 %>% 
  filter(country %in% c("China")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
china_foodplot
```

```{r indiaplot, include=FALSE}
india_foodplot <- final_table2 %>% 
  filter(country %in% c("India")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
india_foodplot
```

```{r Japanplot, include=FALSE}
Japan_foodplot <- final_table2 %>% 
  filter(country %in% c("Japan")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
Japan_foodplot
```

```{r Nigeriaplot, include=FALSE}
Nigeria_foodplot <- final_table2 %>% 
  filter(country %in% c("Nigeria")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
Nigeria_foodplot
```

```{r uschinaplot, include=FALSE}
uschina_foodplot <- final_table2 %>% 
  filter(country %in% c("USA", "China")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
uschina_foodplot
```

```{r usindiaplot, include=FALSE}
usindia_foodplot <- final_table2 %>% 
  filter(country %in% c("USA", "India")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
usindia_foodplot
```

```{r indiachinaplot, include=FALSE}
indiachina_foodplot <- final_table2 %>% 
  filter(country %in% c("India", "China")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
indiachina_foodplot
```

```{r franceusaplot, include=FALSE}
franceus_foodplot <- final_table2 %>% 
  filter(country %in% c("France", "USA")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
franceus_foodplot
```

```{r franceindiaaplot, include=FALSE}
franceindia_foodplot <- final_table2 %>% 
  filter(country %in% c("France", "India")) %>%
  ggplot(aes(x = fct_reorder(food_category, consumption), y = consumption, color = country)) +
  geom_jitter() +
  theme(legend.position = "right") +
  coord_flip()
```

```{r}
franceindia_foodplot
```

```{r uniquecountries, include=FALSE}
unique(final_table2$country)
```

```{r countrynames,  echo=FALSE, warning=FALSE}
knitr::kable(unique(final_table2$country), "html",
             caption = "Country list ",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```


```{r uniquefoodcateghories, include=FALSE}
unique(final_table2$food_category)
```

```{r foodcategory, echo=FALSE, warning=FALSE}

knitr::kable(unique(final_table2$food_category), "html",
             caption = "Food categories within data table",
             booktabs = TRUE,
             longtable = FALSE ) %>%
  kable_styling("striped") 

```



