---
title: 'Car makers industry share  performance '
author: michael
date: '2020-06-07'
slug: car-makers-industry-share-performance
categories:
  - R
tags:
  - Investment research
  - Finance
---

## Trigger to enquiry

Seek to gain sense of relative position and industry performance of Suzuki.   
### Reference

- [sec edgar - financial reports](https://www.sec.gov/edgar/searchedgar/companysearch.html)
- [yahoo finance - share prices](https://finance.yahoo.com/) 
- [Carnegie Mellon R notes](http://www.andrew.cmu.edu/user/achoulde/94842/).

```{r pkgs, include = FALSE}

# Packages

pkgs = c("data.table", "readr", "dplyr", "ggplot2", "tidyverse", "tidyquant", "magrittr", "spelling") 
# data.table package fast writes
# data.table package file loads into data frame
# readr package file loads into tibble 
# ggplot2 package allows qplot and contains the data set mpg
# tidyquant packages gets share prices
# magrittr package extends piping
# spelling package to help in editing document
install.packages(pkgs, repos="https://cran.r-project.org")
inst = lapply(pkgs, library, character.only = TRUE) #load packages
```

### Getting started

Rework Dry Bulk data input sheet for purpose of car industry data screen.  

```{r, include = FALSE}
symbol <- c("7269.T", "7203.T", "7267.T", "7201.T", "F", "GM", "FCA.MI", "VOW.DE", "DAI.DE", "BMW.DE", "UG.PA", "0175.HK", "0489.HK", "600104.SS", "MARUTI.BO", "TATAMOTORS.NS", "005380.KS", "^DJI", "^GSPC", "^BSESN")
company <- c("Suzuki", "Toyota","Honda","Nissan", "Ford", "GM", "Fiat", "VW", "Daimler", "BMW","Peugeot","Geely", "Dongfeng", "SAIC", "Maruti", "Tata", "Hyundai", "Dow Jones Industrial Average", "S&P 500 Index", "Bombay Stock Exchange")
carco <- cbind(symbol, company)
# Save to data file as csv file using data.table package
fwrite(carco, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/carco.csv", row.names = TRUE) # in csv form

carco <- read_csv("/Users/michaelleary/Documents/GitHub/financeScreen/data/carco.csv")

stocks <- c(carco$symbol)
company <- c(carco$company)
#Map a tibble with selected variables companies and index items
stock_list <- tibble(symbol = stocks,
                     company = company)
# companies excluding indexes
stock.list.co <- filter(stock_list, 
                    c(!(symbol=="^DJI" | symbol=="^GSPC" )))

```


```{r, echo = FALSE}

#extract prices
carco.sp <- stock_list %>% tq_get(get = "stock.prices", 
                                  from = "2020-01-01", 
                                  to =Sys.Date())

# In tibble data frame price data for company is stacked on each other.  
```

```{r, include=FALSE}
fwrite(carco.sp, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/carcosp.csv", row.names = TRUE) # in csv form
```

```{r, include = FALSE}
# Row bind in Hyundai share prices

carcosp <- read_csv("/Users/michaelleary/Documents/GitHub/financeScreen/data/carcosp.csv")
carcosp <- carcosp[,-1]
carco.sp <- carcosp

#hyundai <- read_csv("/Users/michaelleary/Documents/GitHub/financeScreen/data/hyundaimotors.csv")

```


```{r, include = FALSE}

#hyundai <- as_tibble(hyundai)

```


```{r, include = FALSE}

#carco.sp2 <- rbind(carco.sp, hyundai)

```

```{r, include=FALSE}
#carco.sp2
```


```{r, include = FALSE}

#list column names

#colnames <- colnames(carco.sp2)
colnames <- colnames(carco.sp)

# prepare to convert character variables symbol, company to factors

#carco.spf <- carco.sp2
carco.spf <- carco.sp

# save character variables
carco.spf %>%
sapply(is.character) %>%
which() %>%
print() ->
chari

# alternate save of character variable names
carco.spf %>%
names() %>%
'['(chari) %>%
print() ->
charc

# convert character variables to factors
carco.spf[charc] %<>% map(factor)  #uncheck
str(carco.spf)
glimpse(carco.spf)

```


```{r, echo=FALSE}
sum.dbulk <- carco.spf %>%
  group_by(symbol, company) %>%
  summarize(curr_price = round(last(adjusted),2), 
            min_price = round(min(adjusted),2), 
            max_price = round(max(adjusted),2),
            pct_from_max = round(100 * (curr_price-max_price)/max_price),
            pct_from_low = round(100 * (curr_price-min_price)/min_price)) %>%
  arrange(desc(pct_from_max))



```

Rank stock price change from 2020 high

```{r}

knitr::kable((sum.dbulk[, c(2,3,5,6,7)]), "markdown", align = "lccccc")

# knitr::kable((sum.dbulk[1:20, 1:7]), "markdown", align = "lccrr")

```

```{r, echo=FALSE}

#print((sum.dbulk[1:20, 1:7]), floating = FALSE)

```

Rank stock price change from 2020 low

```{r, echo=FALSE}
sum.dbulk <- carco.spf %>%
  group_by(symbol, company) %>%
  summarize(curr_price = round(last(adjusted),2), 
            min_price = round(min(adjusted),2), 
            max_price = round(max(adjusted),2),
            pct_from_max = round(100 * (curr_price-max_price)/max_price),
            pct_from_low = round(100 * (curr_price-min_price)/min_price)) %>%
  arrange(desc(pct_from_low))

```

```{r}

knitr::kable((sum.dbulk[, c(2,3,4,6,7)]), "markdown", align = "lccccc")

```

```{r}
# print((sum.dbulk[1:20, 1:7]), floating = FALSE)

```

```{r, echo=FALSE }
# https://ggplot2-book.org/facet.html
adj.plot <- ggplot(data=carco.spf, 
      aes(date, adjusted))
adj.plot + geom_line() + facet_wrap(~symbol, scales = "free")


```

\newpage
# Create index 

```{r, include=FALSE}
# Generate index of components

bse <- carco.spf %>% dplyr::filter((symbol=="^BSESN") )
test <- bse[,c(3,9)]
bsea <- rename(test, bse=adjusted)

dji <- carco.spf %>% dplyr::filter((symbol=="^DJI") )
test <- dji[,c(3,9)]
djia <- rename(test, dji=adjusted)

sp500 <- carco.spf %>% dplyr::filter((symbol=="^GSPC") )
test <- sp500[,c(3,9)]
sp500a <- rename(test, sp500=adjusted)

dst <- dplyr::full_join(bsea, djia, by="date")
head(dst)
dstall <- dplyr::full_join(dst, sp500a, by="date")
head(dstall)
d <- dstall
head(d)

dst <- transform(d, 
                 b.idx=round(100*d$bse/d$bse[1],1), 
                 d.idx=round(100*d$dji/d$dji[1],1), 
                 s.idx=round(100*d$sp500/d$sp500[1],1))
head(dst)

dst <- dst[complete.cases(dst),]

```

```{r, include=FALSE}
# Plot index chart


basketplot <- 
  ggplot(dst, aes(date)) +
  geom_line(aes(y=b.idx, color="Bombay")) + 
  geom_line(aes(y=d.idx, color="DJIA")) +
  geom_line(aes(y=s.idx, color="SP500")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Bombay, DJIA and SP500 index", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplot)
```

# Japan

```{r, include=FALSE}
# Generate index of components

suz <- carco.spf %>% dplyr::filter((symbol=="7269.T") )
test <- suz[,c(3,9)]
suza <- rename(test, suz=adjusted)

toy <- carco.spf %>% dplyr::filter((symbol=="7203.T") )
test <- toy[,c(3,9)]
toya <- rename(test, toy=adjusted)

hon <- carco.spf %>% dplyr::filter((symbol=="7267.T") )
test <- hon[,c(3,9)]
hona <- rename(test, hon=adjusted)

nis <- carco.spf %>% dplyr::filter((symbol=="7201.T") )
test <- nis[,c(3,9)]
nisa <- rename(test, nis=adjusted)


dst1 <- dplyr::full_join(suza, toya, by="date")
head(dst1)
dst2 <- dplyr::full_join(dst1, hona, by="date")
head(dst2)
dst3 <- dplyr::full_join(dst2, nisa, by="date")
head(dst3)

djp <- dst3
head(djp)

djpt <- transform(djp, 
                  suz.idx=round(100*djp$suz/djp$suz[1],1), 
                  toy.idx=round(100*djp$toy/djp$toy[1],1), 
                  hon.idx=round(100*djp$hon/djp$hon[1],1), 
                  nis.idx=round(100*djp$nis/djp$nis[1],1))

head(djpt)

djpt <- djpt[complete.cases(djpt),]

```

```{r, include=FALSE}
# Plot index chart

basketplotjp <- 
  ggplot(djpt, aes(date)) +
  geom_line(aes(y=suz.idx, color="Suzuki")) + 
  geom_line(aes(y=toy.idx, color="Toyota")) +
  geom_line(aes(y=hon.idx, color="Honda")) +
  geom_line(aes(y=nis.idx, color="Nissan")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Japan company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotjp)
```

```{r, include=FALSE}
fwrite(djpt, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/djpt.csv", row.names = TRUE) # in csv form
```

# Europe

```{r, include=FALSE}
# Generate index of components

vw <- carco.spf %>% dplyr::filter((symbol=="VOW.DE") )
test <- vw[,c(3,9)]
vwa <- rename(test, vw=adjusted)

dai <- carco.spf %>% dplyr::filter((symbol=="DAI.DE") )
test <- dai[,c(3,9)]
daia <- rename(test, dai=adjusted)

bmw <- carco.spf %>% dplyr::filter((symbol=="BMW.DE") )
test <- bmw[,c(3,9)]
bmwa <- rename(test, bmw=adjusted)

peu <- carco.spf %>% dplyr::filter((symbol=="UG.PA") )
test <- peu[,c(3,9)]
peua <- rename(test, peu=adjusted)

fca <- carco.spf %>% dplyr::filter((symbol=="FCA.MI") )
test <- fca[,c(3,9)]
fcaa <- rename(test, fca=adjusted)

dst1 <- dplyr::full_join(vwa, daia, by="date")
head(dst1)
dst2 <- dplyr::full_join(dst1, bmwa, by="date")
head(dst2)
dst3 <- dplyr::full_join(dst2, peua, by="date")
head(dst3)
dst4 <- dplyr::full_join(dst3, fcaa, by="date")
head(dst4)


deu <- dst4
head(deu)

deut <- transform(deu, 
                  vw.idx=round(100*deu$vw/deu$vw[1],1), 
                  dai.idx=round(100*deu$dai/deu$dai[1],1), 
                  bmw.idx=round(100*deu$bmw/deu$bmw[1],1),
                  peu.idx=round(100*deu$peu/deu$peu[1],1),
                  fca.idx=round(100*deu$fca/deu$fca[1],1))

head(deut)

deut <- deut[complete.cases(deut),]

```

```{r, include=FALSE}
# Plot index chart

basketploteu <- 
  ggplot(deut, aes(date)) +
  geom_line(aes(y=vw.idx, color="VW")) + 
  geom_line(aes(y=dai.idx, color="Dai")) +
  geom_line(aes(y=bmw.idx, color="BMW")) +
  geom_line(aes(y=peu.idx, color="Peugeot")) +
  geom_line(aes(y=fca.idx, color="Fiat")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Europe company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketploteu)
```

```{r, include=FALSE}
fwrite(deut, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/deut.csv", row.names = TRUE) # in csv form
```


# US

```{r, include=FALSE}
# Generate index of components

fd <- carco.spf %>% dplyr::filter((symbol=="F") )
test <- fd[,c(3,9)]
fda <- rename(test, fd=adjusted)

gm <- carco.spf %>% dplyr::filter((symbol=="GM") )
test <- gm[,c(3,9)]
gma <- rename(test, gm=adjusted)

dst1 <- dplyr::full_join(fda, gma, by="date")
head(dst1)

dus <- dst1
head(dus)

dust <- transform(dus, 
                  fd.idx=round(100*dus$fd/dus$fd[1],1), 
                  gm.idx=round(100*dus$gm/dus$gm[1],1))

head(dust)

dust <- dust[complete.cases(dust),]

```

```{r, include=FALSE}
# Plot index chart

basketplotus <- 
  ggplot(dust, aes(date)) +
  geom_line(aes(y=fd.idx, color="Ford")) + 
  geom_line(aes(y=gm.idx, color="GM")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "US company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotus)
```

```{r, include=FALSE}
fwrite(dust, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/dust.csv", row.names = TRUE) # in csv form
```


# China

```{r, include=FALSE}
# Generate index of components

gee <- carco.spf %>% dplyr::filter((symbol=="0175.HK") )
test <- gee[,c(3,9)]
geea <- rename(test, gee=adjusted)

don <- carco.spf %>% dplyr::filter((symbol=="0489.HK") )
test <- don[,c(3,9)]
dona <- rename(test, don=adjusted)

saic <- carco.spf %>% dplyr::filter((symbol=="600104.SS") )
test <- saic[,c(3,9)]
saica <- rename(test, saic=adjusted)

dst1 <- dplyr::full_join(geea, dona, by="date")
head(dst1)
dst2 <- dplyr::full_join(dst1, saica, by="date")
head(dst2)

dchi <- dst2
head(dchi)

dchit <- transform(dchi, 
                  gee.idx=round(100*dchi$gee/dchi$gee[1],1), 
                  don.idx=round(100*dchi$don/dchi$don[1],1),
                  saic.idx=round(100*dchi$saic/dchi$saic[1],1))

head(dchit)

dchit <- dchit[complete.cases(dchit),]

```

```{r, include=FALSE}
# Plot index chart

basketplotchi <- 
  ggplot(dchit, aes(date)) +
  geom_line(aes(y=gee.idx, color="Geely")) + 
  geom_line(aes(y=don.idx, color="Dongfeng")) +
  geom_line(aes(y=saic.idx, color="SAIC")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "China company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotchi)
```

```{r, include=FALSE}
fwrite(dchit, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/dchit.csv", row.names = TRUE) # in csv form
```


# India

```{r, include=FALSE}
# Generate index of components

mar <- carco.spf %>% dplyr::filter((symbol=="MARUTI.BO") )
test <- mar[,c(3,9)]
mara <- rename(test, mar=adjusted)

tat <- carco.spf %>% dplyr::filter((symbol=="TATAMOTORS.NS") )
test <- tat[,c(3,9)]
tata <- rename(test, tat=adjusted)

dst1 <- dplyr::full_join(mara, tata, by="date")
head(dst1)

dind <- dst1
head(dind)

dindt <- transform(dind, 
                  mar.idx=round(100*dind$mar/dind$mar[1],1), 
                  tat.idx=round(100*dind$tat/dind$tat[1],1))

head(dindt)

dindt <- dindt[complete.cases(dindt),]

```

```{r, include=FALSE}
# Plot index chart

basketplotind <- 
  ggplot(dindt, aes(date)) +
  geom_line(aes(y=mar.idx, color="Maruti Suzuki")) + 
  geom_line(aes(y=tat.idx, color="Tata Motors")) +
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "India company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotind)
```

```{r, include=FALSE}
fwrite(dindt, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/dindt.csv", row.names = TRUE) # in csv form
```

# Korea

```{r, include=FALSE}
# Generate index of components

hyu <- carco.spf %>% dplyr::filter((symbol=="005380.KS") )
test <- hyu[,c(3,9)]
huya <- rename(test, huy=adjusted)

dst1 <- huya
head(dst1)

dkor <- dst1
head(dkor)

dkort <- transform(dkor, 
                  huy.idx=round(100*dkor$huy/dkor$huy[1],1))

head(dkort)

dkort <- dkort[complete.cases(dkort),]

```

```{r, include=FALSE}
# Plot index chart

basketplotkor <- 
  ggplot(dkort, aes(date)) +
  geom_line(aes(y=huy.idx, color="Hyundai")) + 
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Korea company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotkor)
```

```{r, include=FALSE}
fwrite(dkort, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/dkort.csv", row.names = TRUE) # in csv form
```

# Full Index of all itmes

```{r, include=FALSE}
# Generate index of components

dcomb1 <- dplyr::full_join(djp, deu, by="date")
head(dcomb1)
dcomb2 <- dplyr::full_join(dcomb1, dus, by="date")
head(dcomb2)
dcomb3 <- dplyr::full_join(dcomb2, dchi, by="date")
head(dcomb3)
dcomb4 <- dplyr::full_join(dcomb3, dind, by="date")
head(dcomb4)
dcomb5 <- dplyr::full_join(dcomb4, dkor, by="date")
head(dcomb5)

dcomb5t <- transform(dcomb5, 
                  suz.idx=round(100*dcomb5$suz/dcomb5$suz[1],1), 
                  toy.idx=round(100*dcomb5$toy/dcomb5$toy[1], 1),
                  hon.idx=round(100*dcomb5$hon/dcomb5$hon[1],1), 
                  nis.idx=round(100*dcomb5$nis/dcomb5$nis[1], 1),
                  vw.idx=round(100*dcomb5$vw/dcomb5$vw[1],1), 
                  dai.idx=round(100*dcomb5$dai/dcomb5$dai[1],1), 
                  bmw.idx=round(100*dcomb5$bmw/dcomb5$bmw[1],1),
                  peu.idx=round(100*dcomb5$peu/dcomb5$peu[1],1),
                  fca.idx=round(100*dcomb5$fca/dcomb5$fca[1],1),
                 fd.idx=round(100*dcomb5$fd/dcomb5$fd[1], 1),
                  gm.idx=round(100*dcomb5$gm/dcomb5$gm[1],1),
                    gee.idx=round(100*dcomb5$gee/dcomb5$gee[1],1), 
                  don.idx=round(100*dcomb5$don/dcomb5$don[1],1),
                  saic.idx=round(100*dcomb5$saic/dcomb5$saic[1],1),
                    mar.idx=round(100*dcomb5$mar/dcomb5$mar[1], 1),
                  tat.idx=round(100*dcomb5$tat/dcomb5$tat[1],1),
                  huy.idx=round(100*dcomb5$huy/dcomb5$huy[1],1))                    
                     
  
head(dcomb5t)

dcomb5t <- dcomb5t[complete.cases(dcomb5t),]

```

```{r, include=FALSE}
# Plot index chart

basketplotcomb <- 
  ggplot(dcomb5t, aes(date)) +
   geom_line(aes(y=suz.idx, color="Suzuki")) + 
  geom_line(aes(y=toy.idx, color="Toyota")) +
  geom_line(aes(y=hon.idx, color="Honda")) +
  geom_line(aes(y=nis.idx, color="Nissan")) + 
    geom_line(aes(y=vw.idx, color="VW")) + 
  geom_line(aes(y=dai.idx, color="Dai")) +
  geom_line(aes(y=bmw.idx, color="BMW")) +
  geom_line(aes(y=peu.idx, color="Peugeot")) +
  geom_line(aes(y=fca.idx, color="Fiat")) +
  geom_line(aes(y=fd.idx, color="Ford")) + 
  geom_line(aes(y=gm.idx, color="GM")) +
   geom_line(aes(y=gee.idx, color="Geely")) + 
  geom_line(aes(y=don.idx, color="Dongfeng")) +
  geom_line(aes(y=saic.idx, color="SAIC")) +
  geom_line(aes(y=mar.idx, color="Maruti Suzuki")) + 
  geom_line(aes(y=tat.idx, color="Tata Motors")) +
  geom_line(aes(y=huy.idx, color="Hyundai")) + 
  scale_x_date(date_breaks = "months" , date_labels = "%m") +
  labs(title = "Company stock movements 2020", x="Year", y="Index", color="legend") +
  theme_minimal()
  
```

```{r}
plot(basketplotcomb)
```

```{r, include=FALSE}
fwrite(dcomb5t, file = "/Users/michaelleary/Documents/GitHub/financeScreen/data/dcomb5t.csv", row.names = TRUE) # in csv form
```

\newpage
# Index value table

```{r, include=FALSE}
head(dcomb5t)
glimpse(dcomb5t)
index <- dcomb5t[,-c(2:18)]

```

```{r, include=FALSE}

tic <- index

namesC <- c("date", "suz", "toy", "hon", "nis", "vw", "da", "bm", "peu", "fia", "fd", "gm", "gee", "don", "sac", "mar", "tat", "hyu")

names(tic) <- namesC #https://honingds.com/blog/dplyr-rename/

```

```{r}
# https://bookdown.org/yihui/rmarkdown-cookbook/kable.html
knitr::kable(head(tic[,1:17]), "markdown",
             caption = "Index values for car maker share performace in 2020",
             booktabs = TRUE,
             longtable = FALSE )

```

```{r }

knitr::kable(tail(tic[,1:17]), "markdown",
             caption = "Index values for car maker share performace in 2020",
             booktabs = TRUE,
             longtable = FALSE )

```






