---
title: 'Company results template: initial financial prep'
author: Michael
date: '2020-05-27'
slug: company-results-template-initial-financial-prep
categories:
  - R
tags:
  - Finance
  - Investment research
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE)
options(htmltools.dir.version = FALSE)
```

```{r packages, include=FALSE}
# Packages used
pkgs = c("tidyverse", 
         "tidyquant", 
         "magrittr", 
         "data.table", 
         "fBasics", 
         "gdata",
         "dplyr",
         "ggplot2",
         "spelling") 
#package names: 
#tidyverse for operations using dplyr's piping feature; 
#magrittr %<>% function pipes left hand data to right side and returns result to left hand overwriting original contents on left hand
#data.table for fast csv save
#gdata has keep function for permitting convenient trim of env objects with sure=TRUE

install.packages(pkgs, repos="https://cran.r-project.org")

inst = lapply(pkgs, library, character.only = TRUE) #load packages

```

# Financials

Suzuki Motors (7269.T, JPY3700) announced year end FY2019 (April 2019 through March 2020) results that show a decline in net sales and earnings. In the following articles will examine.   

1.    headline earnings; 
2.    revenue and operating profit mix; 
3.    balance sheet position; 
4.    cash flow generation and use; 
5.    ratio analysis; 
6.    consensus earnings outlook; 
7.    valuation.  

But for now will assemble input data that will then wrangle into a form suitable for future analysis and presentation. 

```{r topchunkdata, include=FALSE}
topchunk <- read.csv("/Users/michaelleary/Dropbox/Rwork/michaelLeary-io-site/data/topchunk.csv", stringsAsFactors = TRUE)

head(topchunk)

```

```{r topchunkdataset, include=FALSE}
# examine variables and variable types
class(topchunk)
glimpse(topchunk)
str(topchunk)
topchunkcol <- colnames(topchunk)

```

```{r columnset, include=FALSE}
top.col <- topchunk[,c("period", 
                     "net_sales",
                     "op_inc",
                     "ord_inc",
                     "net_inc",
                     "div_amt",
                     "avg_sharesiss",
                     "tot_ass",
                     "share_eq",
                     "liquid_ass",
                     "tot_debt",
                     "ppe",
                     "cost_sales",
                     "sga",
                     "car_sales",
                     "bike_sales",
                     "marine_sales",
                     "car_opinc",
                     "bike_opinc",
                     "marine_opinc",
                     "jp_sales",
                     "india_sales",
                     "other_sales",
                     "jp_ppe",
                     "india_ppe",
                     "others_ppe"
                     )]

```

```{r calculatedvariables, include=FALSE}
# items to be computed and added to input data include:
# eps
# dps
# roe
# bvps
# net debt / sh.equity
# dupont analysis 
# margins

fullset.topcol <- mutate(top.col,
                    eps = round((net_inc / avg_sharesiss),1), 
                    dps = round((div_amt / avg_sharesiss),1),
                    bvps = round((share_eq / avg_sharesiss),),
                    roe = round(100 * (net_inc / share_eq),2),
                    assetturnover = round((net_sales / tot_ass),2),
                    leverage = round((tot_ass / share_eq),2),
                    netmargin = round((net_inc / net_sales),2),
                    dupont_roe = round(100 * (netmargin * assetturnover * leverage),2))

# company net debt

codebt <- fullset.topcol[,c("period", "tot_debt")]
co_debt <- mutate(codebt)
rownames(co_debt) <- co_debt$period
co_debt$period <- NULL
co_debt_transpose <- as.data.frame(t(as.matrix(co_debt)))

# company liquidity

coliquid <- fullset.topcol[,c("period", "liquid_ass")]
co_liquid <- mutate(coliquid)
rownames(co_liquid) <- co_liquid$period
co_liquid$period <- NULL
co_liquid_transpose <- as.data.frame(t(as.matrix(co_liquid)))

# company net debt mutated onto fullset.top

fs.top <- mutate(fullset.topcol,
                    debt = co_debt$tot_debt,
                    liquidtassets = co_liquid$liquid_ass,  
                    net_debt = debt - liquidtassets,
                    Pct_Net_D_of_E = round(100*(net_debt/share_eq),))

```


```{r fullsettranspose, include=FALSE}
# Dataset includes computed items and d/e and transpose and added column for yoy change

tvultable <- fs.top[c(1:38)]

rownames(tvultable) <- tvultable$period
tvultable$period <- NULL

t_transpose <- as.data.frame(t(as.matrix(tvultable)))

fullset <- setDT(t_transpose, keep.rownames = TRUE)[]

# rename auto generated column name with previously saved label

names(fullset)[1] <- "Item"

fs.yoy <- fullset %>%
  mutate(abschange = round((fullset$FY2019 - fullset$FY2018),),
         pct = round(100 * (abschange / fullset$FY2018),1)) %>%
  select("Item", "FY2019", "FY2018", "pct", "abschange")

```


```{r fullsettable, echo=FALSE}

fs.yoy

```

